---
title: "CLL drug combinations ex-vivo: 10x10 screens"
author: "Britta Velten"
date: "4 September 2018"
output:
  BiocStyle::html_document:
    toc: true
    toc_depth: 3
params: 
    today: 180914
---

#Introduction

Takes raw data files for 10x10 screen.

```{r, echo=F}
library(knitr )
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
options(stringAsFactors = FALSE)
```

```{r, echo=F}
library(ggplot2)
library(RColorBrewer)
library(Biobase)
library(abind)
library(grid)
library(gtable)
library(reshape2)
library(gridExtra)
require(pracma)
library(reshape2)
library(dplyr)
library(pheatmap)
library(magrittr)
library(lattice)
library(tidyverse)
library(xlsx)
source("plot_utils.R")
```

```{r}
# setwd("~/Documents/cll/MarinaDrugComb/DrugCombi_Code4Ms/")
# datadir <- "~/Documents/cll/MarinaDrugComb/rawData"
datadir <- "/Volumes/huber/projects/nct/cll/RawData/DrugScreens/Marina_CombiScreen/"
outdir = "out"
today <- params$today
figdir = paste0("figs", today, "/figures10x10/")
if(!dir.exists(figdir)) dir.create(figdir)
knitr::opts_chunk$set(dev = c("png", "pdf"), fig.path = paste0("figs", today, "/figures10x10/"))
```


```{r}
load(file.path(outdir, paste0("MarinaCombiDataAfterQC_",today,".RData")))
datadir10x10 <- file.path(datadir,"10x10")
```


# Get plate set-up
From Prtocol get set-up of the plate (i.e. which substance in which concentration in which well)
```{r}
FileSetUp10x10 <- as.matrix(read.xlsx(file.path(datadir10x10,"Protokoll 10 x 10 Validierung Drugscreen2.xlsx"),2))

# get set-up for base/test drugs
  tableIdxBase <- which(FileSetUp10x10=="A", arr.ind = T)[1,]
  SetUpBase <- FileSetUp10x10[tableIdxBase[1]+0:15,tableIdxBase[2]+1:24]
  rownames(SetUpBase) <- FileSetUp10x10[tableIdxBase[1]+0:15,tableIdxBase[2]]
  colnames(SetUpBase) <- FileSetUp10x10[tableIdxBase[1]-1,tableIdxBase[2]+1:24]
  SetUpBase <- rownames_to_column(as.data.frame(SetUpBase), var = "WellLetter") %>%
    gather(key="WellNo", value = "base", -WellLetter)
  
  # get concentration values per column of plate
  conc_base_df <- tibble(WellNo = FileSetUp10x10[tableIdxBase[1]-1,tableIdxBase[2]+1:24],
                      base_conc = FileSetUp10x10[tableIdxBase[1]-2,tableIdxBase[2]+1:24])
  
  SetUpBase %<>% left_join(conc_base_df, by = "WellNo")
  SetUpBase %<>% mutate(base_conc = ifelse(base=="DM", NA, base_conc))
  SetUpBase <- filter(SetUpBase, !is.na(base))
  
  
# get set-up for ibrutinib
  tableIdxIbr <- which(FileSetUp10x10=="A", arr.ind = T)[2,]
  SetUpIbr <- FileSetUp10x10[tableIdxIbr[1]+0:15,tableIdxIbr[2]+1:24]
  rownames(SetUpIbr) <- FileSetUp10x10[tableIdxIbr[1]+0:15,tableIdxIbr[2]]
  colnames(SetUpIbr) <- FileSetUp10x10[tableIdxIbr[1]-1,tableIdxIbr[2]+1:24]
  SetUpIbr <- rownames_to_column(as.data.frame(SetUpIbr), var = "WellLetter") %>%
    gather(key="WellNo", value = "combi", -WellLetter)
  
  # get concentration values per row of plate
  conc_combi_df <- tibble(WellLetter = FileSetUp10x10[tableIdxIbr[1]+0:15,tableIdxIbr[2]],
                          combi_conc = FileSetUp10x10[tableIdxIbr[1]+(0:15),tableIdxIbr[2]+25])
  conc_combi_df %<>% mutate(combi_conc = ifelse(combi_conc == "Concentration (Synergie-Substanz)", NA, combi_conc))

  SetUpIbr %<>% left_join(conc_combi_df, by = "WellLetter")
  
  # set conc for DM to NA
  SetUpIbr %<>% mutate(combi_conc = ifelse(!is.na(combi), ifelse(combi=="DM", NA, combi_conc), combi_conc))

  #fill in NAs with a conc value by Ibrutinib
  SetUpIbr %<>% mutate(combi = ifelse(combi_conc %in% paste0("c", 1:10), "Ibrutinib", combi))
  SetUpIbr <- filter(SetUpIbr, !is.na(combi))
                       
  # fill in row with changing values
  SetUpIbr %<>% mutate(combi_conc = ifelse(grepl("Ibru c",combi), sub("Ibru ", "", combi), combi_conc))
  SetUpIbr %<>% mutate(combi = ifelse(grepl("Ibru c",combi), "Ibrutinib", combi))

  # merge 
  SetUp <- full_join(SetUpIbr, SetUpBase, by=c("WellLetter", "WellNo"))
  SetUp %<>% mutate(WellLetter = tolower(WellLetter))
```


# Read in values from 10x10 plates
```{r}
import10x10Data <- function(filename, SetUp){
  print(filename)
  data <- read.xlsx2(file.path(datadir10x10,filename),1, header = F, startRow=3)
  
  #info from filename
  NameComp <- strsplit(filename, "_")[[1]]
  patID <- NameComp[grep("P0",NameComp)] 
  
  # in files named PatientID.... first named drug is second in screen
  # in files named 10x10 Validierung..... first named drug is second in screen.
  if(grepl("^10x10 Validierung", filename)){
  TestDrugA <- NameComp[3] 
  TestDrugB <- sub(".xlsx","", NameComp[4])
  } else {
      TestDrugB <- NameComp[3] 
      TestDrugA <- sub(".xlsx","", NameComp[4])
  }
    
  
  #table from excel file (ignore meta data about screen)
  tableIdx <- which(data=="<>", arr.ind = T)
  stopifnot(nrow(tableIdx)==1)
  plateValues <- data[tableIdx[1]+1:11,tableIdx[2]+1:24]
  colnames(plateValues) <- 1:ncol(plateValues)
  rownames(plateValues) <- letters[2 + 1:nrow(plateValues)]  #labels start with C
  
  # turn into data.frame
  df <- rownames_to_column(plateValues, var = "WellLetter")
  df %<>% gather(key="WellNo", value="rawValue", -WellLetter)
  df %<>% mutate(rawValue = as.numeric(rawValue))
  
  # add setup info
  df %<>% left_join(SetUp,  by=c("WellLetter", "WellNo"))
  
  # fill in test drug names: test drug 1 is first of file name
  df %<>% mutate(base = ifelse(base =="DM", "DM", ifelse(base==1,TestDrugA, TestDrugB)))
  df %<>% rename(BaseDrug = base)
  df %<>% rename(CombiDrug = combi)

  # normalized by pure DMSO median
  dfDMSO <- filter(df, BaseDrug=="DM" , CombiDrug=="DM")
  df %<>% mutate(pureDMSOMedian = median(dfDMSO$rawValue))
  df %<>% mutate(pureDMSOMean = mean(dfDMSO$rawValue))
  df %<>% mutate(normalizedValue = rawValue/pureDMSOMedian)

  # get table of concentrations (for Ibrutinib, drug A and drug B)
  # (bug in reaad.xlsx but dilution factor is always 2 and c10 is zero, replace by hand)
  conctableIdx <- which(data=="µM"| data=="nM", arr.ind = T)
  concArray <- sapply(1:3, function(i) {
    concvec <- rep(0,10)
    names(concvec) <- paste("c", 1:10, sep="")
    concvec[1] <- as.numeric(as.character(data[conctableIdx[i,1]+0,conctableIdx[i,2]+1]))
    if(data[conctableIdx[i,1],conctableIdx[i,2]]=="nM")   concvec[1] <- concvec[1]/1000
    for(j in 2:9) concvec[j] <- concvec[j-1]/2
    concvec
  })
  colnames(concArray) <- data[conctableIdx-c(1,1)]
  concArray <- cbind(concArray,DM=rep(NA,10))
  colnames(concArray)[colnames(concArray)=="CAL-101"] <- "CAL101"
  colnames(concArray)[colnames(concArray)=="Abt-199"] <- "Abt199"
  colnames(concArray)[colnames(concArray)=="ABT-263"] <- "ABT263"
  concValueDf <- concArray %>% as.data.frame() %>% 
    rownames_to_column(var = "concId") %>% gather(key="drug", value = "concValue", -concId)
  
  # add true conc values to df
  df %<>% left_join(concValueDf, by = c("CombiDrug" = "drug", "combi_conc" = "concId"))
  df %<>% rename(concCvalue = concValue)

  df %<>% left_join(concValueDf, by = c("BaseDrug" = "drug", "base_conc" = "concId"))
  df %<>% rename(concBvalue = concValue)
  df %<>% mutate(PatientID = patID)

  return(df)
  }
```

```{r}
allFiles <- setdiff(list.files(file.path(datadir10x10), "*.xlsx"), list.files(file.path(datadir10x10),"Protokoll"))
listOfTables <- lapply(allFiles, function(file) import10x10Data(file, SetUp))
for(l in 1:length(listOfTables)) listOfTables[[l]]$PlateID <- sprintf("10x10PL%03d", l)
df10x10 <- do.call(rbind,listOfTables)
```

```{r}
# use nice drug names
df10x10 %<>% mutate(BaseDrugName = ifelse(BaseDrug == "CAL101", "Idelalisib", 
                           ifelse(BaseDrug == "Abt199", "Venetoclax",
                                  ifelse(BaseDrug == "ABT263","Navitoclax", BaseDrug))))
```

## QC
Check whether the concentratino in the xls files are plausible: They are different in each experiment, consistent would be 2µM for navitoclax and venetoclax and 20µM for afatinib, idelalisib and ibrutinib in 2-fold serial dilutions, ending at 0.
```{r}
conc_BCL2 <- sapply(0:9, function(i) 2 * 0.5^i)
conc_BCL2[10] <- 0
names(conc_BCL2) <- paste0("c", 1:10)
conc_other <- sapply(0:9, function(i) 20 * 0.5^i)
conc_other[10] <- 0
names(conc_other) <- paste0("c", 1:10)
df10x10corr <- mutate(df10x10, concBvalue = ifelse(BaseDrugName %in% c("Afatinib", "Idelalisib"), conc_other[base_conc],
                                 ifelse(BaseDrugName %in% c("Navitoclax", "Venetoclax"),  conc_BCL2[base_conc], NA)))
df10x10corr <-mutate(df10x10corr, concCvalue = ifelse(CombiDrug == "Ibrutinib", conc_other[combi_conc], NA))
```

Check that concentration values for B are correct (different per patient) in the spreadsheeds. YES
```{r, fig.width=10}
gg1 <- filter(df10x10, concCvalue == 0, CombiDrug == "Ibrutinib", BaseDrugName!="DM") %>%
  ggplot(aes(x= concBvalue, y=normalizedValue, col= PatientID)) +
  geom_line() + facet_wrap(~BaseDrugName, scales = "free_x")+ggtitle("concentrations as spreadsheets")

gg2 <- filter(df10x10corr, concCvalue == 0, CombiDrug == "Ibrutinib", BaseDrugName!="DM") %>%
  ggplot(aes(x= concBvalue, y=normalizedValue, col= PatientID)) +
  geom_line() + facet_wrap(~BaseDrugName, scales = "free_x") +ggtitle("concentrations as in manuscript")

cowplot::plot_grid(gg1,gg2, ncol=2)

filter(df10x10,  CombiDrug == "Ibrutinib", BaseDrugName!="DM") %>%
  ggplot(aes(x= concBvalue, y=normalizedValue, col= PatientID)) +
  geom_line() + facet_grid(concCvalue ~BaseDrugName, scales = "free_x")
```

## Overview
```{r}
# data overview
df10x10 %>% group_by(BaseDrugName, CombiDrug) %>%
  summarize(nPats = length(unique(PatientID)), pats = paste(unique(PatientID), collapse =", "))

df10x10 %>% filter(base_conc== "c1") %>%
  group_by(BaseDrugName, concBvalue) %>%
  summarize(nP = length(unique(PatientID)))
```

## Data export
```{r}
save(df10x10, file = file.path(outdir, paste0("df10x10_",today,".Rdata")))

if(!dir.exists(file.path(outdir, "tables"))) dir.create(file.path(outdir, "tables"))
write.csv(df10x10, file= paste0(outdir,"/tables/10x10_",today,".csv"))
```


# Tile-Plots of normalized viabilties

```{r}
knitr::opts_chunk$set(fig.width=8)
```

## Navitoclax
```{r 10x10_Navitoclax_P0016}
plotTiles10x10(df10x10, "Navitoclax", "P0016")
```
```{r 10x10_Navitoclax_P0347}
plotTiles10x10(df10x10, "Navitoclax", "P0347")
```
```{r 10x10_Navitoclax_P0445}
plotTiles10x10(df10x10, "Navitoclax", "P0445")
```
```{r 10x10_Navitoclax_P0309}
plotTiles10x10(df10x10, "Navitoclax", "P0309")
```
```{r 10x10_Navitoclax_P0369}
plotTiles10x10(df10x10, "Navitoclax", "P0369")
```

## Afatinib
```{r 10x10_Afatinib_P0010}
plotTiles10x10(df10x10, "Afatinib", "P0010")
```
```{r 10x10_Afatinib_P0050}
plotTiles10x10(df10x10, "Afatinib", "P0050")
```
```{r 10x10_Afatinib_P0445}
plotTiles10x10(df10x10, "Afatinib", "P0069")
```
```{r 10x10_Afatinib_P0480}
plotTiles10x10(df10x10, "Afatinib", "P0480")
```
```{r 10x10_Afatinib_P0484}
plotTiles10x10(df10x10, "Afatinib", "P0484")
```

## Venetoclax
```{r 10x10_Venetoclax_P0016}
plotTiles10x10(df10x10, "Venetoclax", "P0016")
```
```{r 10x10_Venetoclax_P0347}
plotTiles10x10(df10x10, "Venetoclax", "P0347")
```
```{r 10x10_Venetoclax_P0445}
plotTiles10x10(df10x10, "Venetoclax", "P0445")
```

# Assesing synergy using SynergyFinder
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4759128/
```{r}
library("synergyfinder")
```

## Single-Dose Response Curves
```{r}
ggplot(filter(df10x10, BaseDrugName == "Navitoclax", CombiDrug == "Ibrutinib"),
       aes(x = (concBvalue), y= normalizedValue, col = PatientID, shape = PatientID %in% c("P0369", "P0309"))) +
  geom_point() + facet_wrap(~concCvalue) +scale_x_log10() 

ggplot(filter(df10x10, BaseDrugName == "Navitoclax", CombiDrug == "Ibrutinib"),
       aes(x = (concCvalue), y= normalizedValue, col = PatientID, shape = PatientID %in% c("P0369", "P0309"))) +
  geom_point() + facet_wrap(~(concBvalue*1000)) +scale_x_log10() 
```

## Dose Response curves of combination drug
```{r}
# B drug curves
ggplot(filter(df10x10, BaseDrugName == "Navitoclax", CombiDrug == "Ibrutinib", concCvalue==0),
       aes(x = (concBvalue), y= normalizedValue, col = PatientID, shape = PatientID %in% c("P0369", "P0309"))) +
  geom_point() + facet_wrap(~(concCvalue*1000)) +scale_x_log10() +geom_line()
ggplot(filter(df10x10, BaseDrugName == "Venetoclax", CombiDrug == "Ibrutinib", concCvalue==0),
       aes(x = (concBvalue), y= normalizedValue, col = PatientID, shape = PatientID %in% c("P0369", "P0309"))) +
  geom_point() + facet_wrap(~(concCvalue*1000)) +scale_x_log10() +geom_line()
ggplot(filter(df10x10, BaseDrugName == "Afatinib", CombiDrug == "Ibrutinib", concCvalue==0),
       aes(x = (concBvalue), y= normalizedValue, col = PatientID, shape = PatientID %in% c("P0369", "P0309"))) +
  geom_point() + facet_wrap(~(concCvalue*1000)) +scale_x_log10() +geom_line()
```

## Dose Response curves of Ibrutinib
```{r}
# C Drug curves
ggplot(filter(df10x10, BaseDrugName == "Navitoclax", CombiDrug == "Ibrutinib", concBvalue==0),
       aes(x = (concCvalue), y= normalizedValue, col = PatientID, shape = PatientID %in% c("P0369", "P0309"))) +
  geom_point() + facet_wrap(~(concBvalue*1000)) +scale_x_log10() +geom_line()

ggplot(filter(df10x10, BaseDrugName == "Venetoclax", CombiDrug == "Ibrutinib", concBvalue==0),
       aes(x = (concCvalue), y= normalizedValue, col = PatientID, shape = PatientID %in% c("P0369", "P0309"))) +
  geom_point() + facet_wrap(~(concBvalue*1000)) +scale_x_log10() +geom_line()

ggplot(filter(df10x10, BaseDrugName == "Afatinib", CombiDrug == "Ibrutinib", concBvalue==0),
       aes(x = (concCvalue), y= normalizedValue, col = PatientID, shape = PatientID %in% c("P0369", "P0309"))) +
  geom_point() + facet_wrap(~(concBvalue*1000)) +scale_x_log10() +geom_line()
```
The drug curves of Ibrutinib alone are strange and only in the last concentration something happens. Thereby the Loewe synergy cannot be well assessed as no conc with similar effect as Ibrutinib and lowest conc of drug B can be found.


## Loewe score
```{r Loewe_surface, fig.width=10, warning= FALSE}
dfLoewe <- plotCITiles(df10x10, "Loewe")
```

Notes: Changinng the concentration of Navitoclax by a factor of 10 leads to a different Loewe synergy, the ZIP index remains the same. Dose response curves and matrices as obtained from PlotDoseResponse are not affected  by the change.
This is due to calculating the synergy as response - maximal response on sum of both concentration in one of the two single fits, if the Loewe equation cannot be colved. MyLoewe takes for these cases  NA. For details see synergyfinder_problems1/2. (altenrative: the clostest value in 0 to 100 to fulfil this equation, can produce high synergy at high concentrations, where effect is clost to 100)

## Loewe Score with missing parts
```{r myLoewe_surface}
source("~/Documents/CLL/MarinaDrugComb/DrugCombi_Code4Ms/myLoewe.R")
dfmyLoewe <- plotCITiles(df10x10, "myLoewe", cutoff = 100)
```

## ZIP score
```{r ZIP_surface}
dfZIP <- plotCITiles(df10x10, "ZIP")
```

# Summarize scores
## Navitoclax
### Loewe
```{r summary_Navitoclax_Loewe, fig.width = 10 }
# Strong antagonism for low concentrations? Is ZIP more stable for small effects?
ggNavi <- plotSummaryLoewe(dfLoewe = dfLoewe, "Navitoclax", type = "row")
ggNavi
```

### Loewe Score with missing parts
```{r summary_Navitoclax_myLoewe, fig.width = 10 }
# Strong antagonism for low concentrations? Is ZIP more stable for small effects?
ggNavi <- plotSummaryLoewe(dfLoewe = dfmyLoewe, "Navitoclax", type = "row")
ggNavi
```

### ZIP score
```{r summary_Navitoclax_ZIP, fig.width = 10 }
# Strong antagonism for low concentrations? Is ZIP more stable for small effects?
ggNavi <- plotSummaryLoewe(dfLoewe = dfZIP, "Navitoclax", type = "row")
ggNavi
```

## Afatinib
### Loewe
```{r summary_Afatinib_Loewe, fig.width = 10 }
ggAfa <- plotSummaryLoewe(dfLoewe = dfLoewe, "Afatinib", type ="row")
ggAfa
```

### Loewe Score with missing parts
```{r summary_Afatinib_myLoewe, fig.width = 10 }
ggAfa <- plotSummaryLoewe(dfLoewe = dfmyLoewe, "Afatinib", type ="row")
ggAfa
```

### ZIP score
```{r summary_Afatinib_ZIP, fig.width = 10 }
ggAfa <- plotSummaryLoewe(dfLoewe = dfZIP, "Afatinib", type ="row")
ggAfa
```

## Venetoclax
### Loewe
```{r summary_Venetoclax_Loewe, fig.width = 10 }
ggVen <- plotSummaryLoewe(dfLoewe = dfLoewe, "Venetoclax", type ="row")
ggVen
```

### Loewe Score with missing parts
```{r summary_Venetoclax_myLoewe, fig.width = 10 }
ggVen <- plotSummaryLoewe(dfLoewe = dfmyLoewe, "Venetoclax", type ="row")
ggVen
```

### ZIP score
```{r summary_Venetoclax_ZIP, fig.width = 10 }
ggVen <- plotSummaryLoewe(dfLoewe = dfZIP, "Venetoclax", type ="row")
ggVen
```

## Idelalisib
### Loewe
```{r summary_Idelalisib_Loewe, fig.height = 8 }
ggIdel <- plotSummaryLoewe(dfLoewe = dfLoewe, "Idelalisib")
ggIdel
```
### Loewe Score with missing parts
```{r summary_Idelalisib_myLoewe, fig.height = 8 }
ggIdel <- plotSummaryLoewe(dfLoewe = dfmyLoewe, "Idelalisib")
ggIdel
```
### ZIP score
```{r summary_Idelalisib_ZIP, fig.height = 8 }
ggIdel <- plotSummaryLoewe(dfLoewe = dfZIP, "Idelalisib")
ggIdel
```


```{r}
sessionInfo()
```

# Understanding scores and strong antagonism
How to adjust for differences in concentrations? D
ifferent factors lead to different results. Why?
```{r}
df10x10mod <- df10x10
df10x10mod$concBvalue <- 10*df10x10mod$concBvalue  # makes concentration ranges similar to Ibrutinib for Navitoclax
dfLoewemod <- plotCITiles(df10x10mod, "Loewe")

plotSummaryLoewe(dfLoewe = dfLoewemod, "Navitoclax", type = "row")
plotSummaryLoewe(dfLoewe = dfLoewe, "Navitoclax", type = "row")
```

ZIP model is not influenced by this.
```{r}
dfZIPmod <- plotCITiles(df10x10mod, "ZIP")
plotSummaryLoewe(dfLoewe = dfZIPmod, "Navitoclax", type = "row")
plotSummaryLoewe(dfLoewe = dfZIP, "Navitoclax", type = "row")
```

<!-- # Comparison -->
<!-- Compare to other scores -->
<!-- ```{r score_comparison, fig.height=5} -->
<!-- Bdrs <- unique(df10x10$BaseDrugName) -->
<!-- Bdrs  <- Bdrs[Bdrs != "DM"] -->
<!-- df_scores <- lapply(c("Loewe", "ZIP", "HSA", "Bliss"), function(CI_type) { -->
<!--  lapply(Bdrs, function(dr){ -->
<!--     data <- filter(df10x10, BaseDrugName == dr, CombiDrug == "Ibrutinib") -->
<!--    lapply(unique(data$PatientID), function(pat){ -->
<!--      dfpat <- filter(data, PatientID == pat) %>% -->
<!--   select(Replicate = PatientID, DrugRow = BaseDrugName, DrugCol = CombiDrug, -->
<!--          ConcRow = concBvalue, ConcCol = concCvalue, Response = normalizedValue, -->
<!--          Row = base_conc, Col = combi_conc) %>% -->
<!--   mutate(BlockID = 1, ConcRowUnit = "μM", ConcColUnit = "μM")   %>% # only one drug-drug combination -->
<!--   mutate(Row = sub("c", "", Row), -->
<!--          Col = sub("c", "", Col), -->
<!--          Response = 100 * Response) # need percentage -->
<!-- dfpat <- ReshapeData(dfpat, data.type = "viability") -->
<!-- dfpat$dose.response.mats[[1]] <- dfpat$dose.response.mats[[1]][order(as.numeric(rownames(dfpat$dose.response.mats[[1]]))), order(as.numeric(colnames(dfpat$dose.response.mats[[1]])))] -->
<!-- synergy.score <- CalculateSynergy(dfpat,method = CI_type, correction = TRUE, Emin = 0, Emax = 100, nan.handle = "LL4") -->
<!-- data.frame(pat=pat, drB =dr, score = mean(synergy.score$scores[[1]]), method = CI_type) -->
<!--    }) %>% bind_rows() -->
<!--  }) %>% bind_rows() -->
<!-- }) %>% bind_rows() -->

<!-- # compare to effect based synergies (needs running CombiSynergy.Rmd first) -->
<!-- if(file.exists(file.path(outdir, paste0("MarinaCombiSyns_",today,".RData")))){ -->
<!-- load("out/MarinaCombiSyns_",today,".RData") -->
<!-- dfEffectBased <- dfsynSummaryConc %>% filter(CDrugAbrv == "Ibrutinib (100nM)") %>% -->
<!--   select(BDrugName, addModelSImean, hsaCImean, PatientID)%>% -->
<!--   gather(key = "method", value="score", -c("PatientID","BDrugName")) %>% rename(pat = PatientID, drB = BDrugName) %>% filter(pat %in% df_scores$pat, drB %in% df_scores$drB) -->

<!-- df_scores %<>% mutate(method = paste0(method, "_10x10")) -->
<!-- df4plot <- rbind(df_scores,dfEffectBased) -->
<!-- } else df4plot <- df_scores -->

<!-- df4plot %>% ggplot(aes(x=pat, y=score, fill=pat)) +  -->
<!--   facet_grid(method~drB, scales = "free") + geom_bar(stat="identity") +  -->
<!--   theme(axis.text.x = element_blank()) + -->
<!--   guides(fill =F) + scale_fill_manual(values=patcol) +  -->
<!--   geom_hline(yintercept = 0) -->
<!-- ``` -->



<!-- # Isobologram - Analysis -->
<!-- Alternative to Synergyfinder -->
<!-- ```{r} -->
<!-- library(BIGL) -->
<!-- evalIsobols <- FALSE -->
<!-- ``` -->

<!-- ```{r isobologram_Venetoclax, eval = evalIsobols} -->
<!-- data <- filter(df10x10, BaseDrugName == "Venetoclax") %>% -->
<!--   select(experiment = PatientID, cpd1 = BaseDrugName, cpd2 = CombiDrug, d1 = concBvalue, d2 = concCvalue, effect = normalizedValue) -->
<!-- marginalFit <- fitMarginals(data, method = "optim", names = c("cpd1", "cpd2")) -->
<!-- summary(marginalFit) -->
<!-- rs <- fitSurface(data, marginalFit, -->
<!--                  null_model = "loewe", -->
<!--                  B.CP = 50, statistic = "meanR", parallel = FALSE) -->
<!-- isobologram(rs) -->
<!-- summary(rs) -->

<!-- rs <- fitSurface(data, marginalFit, -->
<!--                  null_model = "hsa", -->
<!--                  B.CP = 50, statistic = "meanR", parallel = FALSE) -->
<!-- isobologram(rs) -->
<!-- summary(rs) -->
<!-- ``` -->

<!-- ```{r isobologram_Navitoclax, eval = evalIsobols} -->
<!-- data <- filter(df10x10, BaseDrugName == "Navitoclax", CombiDrug == "Ibrutinib") %>% -->
<!--   select(experiment = PatientID, cpd1 = BaseDrugName, cpd2 = CombiDrug, d1 = concBvalue, d2 = concCvalue, effect = normalizedValue) -->
<!-- marginalFit <- fitMarginals(data, method = "optim", names = c("cpd1", "cpd2")) -->
<!-- summary(marginalFit) -->
<!-- rs <- fitSurface(data, marginalFit, -->
<!--                  null_model = "loewe", -->
<!--                  B.CP = 50, statistic = "meanR", parallel = FALSE) -->
<!-- isobologram(rs) -->
<!-- summary(rs) -->

<!-- rs <- fitSurface(data, marginalFit, -->
<!--                  null_model = "hsa", -->
<!--                  B.CP = 50, statistic = "meanR", parallel = FALSE) -->
<!-- isobologram(rs) -->
<!-- summary(rs) -->
<!-- ``` -->

<!-- ```{r isobologram_Afatinib, eval = evalIsobols} -->
<!-- data <- filter(df10x10, BaseDrugName == "Afatinib") %>% -->
<!--   select(experiment = PatientID, cpd1 = BaseDrugName, cpd2 = CombiDrug, d1 = concBvalue, d2 = concCvalue, effect = normalizedValue) -->
<!-- marginalFit <- fitMarginals(data, method = "optim", names = c("cpd1", "cpd2")) -->
<!-- summary(marginalFit) -->
<!-- rs <- fitSurface(data, marginalFit, -->
<!--                  null_model = "loewe", -->
<!--                  B.CP = 50, statistic = "meanR", parallel = FALSE) -->
<!-- isobologram(rs) -->
<!-- summary(rs) -->

<!-- rs <- fitSurface(data, marginalFit, -->
<!--                  null_model = "hsa", -->
<!--                  B.CP = 50, statistic = "meanR", parallel = FALSE) -->
<!-- isobologram(rs) -->
<!-- summary(rs) -->
<!-- ``` -->

<!-- ```{r isobologram_Idelalisib, eval = evalIsobols} -->
<!-- data <- filter(df10x10, BaseDrugName == "Idelalisib") %>% -->
<!--   select(experiment = PatientID, cpd1 = BaseDrugName, cpd2 = CombiDrug, d1 = concBvalue, d2 = concCvalue, effect = normalizedValue) -->
<!-- marginalFit <- fitMarginals(data, method = "optim", names = c("cpd1", "cpd2")) -->
<!-- summary(marginalFit) -->
<!-- rs <- fitSurface(data, marginalFit, -->
<!--                  null_model = "loewe", -->
<!--                  B.CP = 50, statistic = "meanR", parallel = FALSE) -->
<!-- isobologram(rs) -->
<!-- summary(rs) -->

<!-- rs <- fitSurface(data, marginalFit, -->
<!--                  null_model = "hsa", -->
<!--                  B.CP = 50, statistic = "meanR", parallel = FALSE) -->
<!-- isobologram(rs) -->
<!-- summary(rs) -->
<!-- ``` -->



