---
title: "CLL drug combinations ex-vivo: Find synergistic effect of drug-drug combinations"
author: "Britta Velten"
date: "2 Sept 2018"
output:
  BiocStyle::html_document:
    toc: true
    toc_depth: 3
params: 
    today: 180905
---
#Introduction

This file analyses the data from the combination screen for combination with a second compound (drugC).

```{r, echo=F}
library(knitr )
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

```{r, echo=F}
library(dplyr)
library(RColorBrewer)
library(Biobase)
library(abind)
library(grid)
library(gtable)
library(reshape2)
library(gridExtra)
require(pracma)
library(reshape2)
library(pheatmap)
library(magrittr)
library(lattice)
library(tidyverse)
source("plot_utils.R")
```

```{r, echo=F}
# setwd("~/Documents/cll/MarinaDrugComb/DrugCombi_Code4Ms/")
# datadir <- "~/Documents/cll/MarinaDrugComb/rawData"
datadir <- "/Volumes/huber/projects/nct/cll/RawData/DrugScreens/Marina_CombiScreen/"
outdir <- "out"
today <- params$today
figdir <- paste0("figuresCombiSynergy", today, "/")
if(!dir.exists(figdir)) dir.create(figdir)
knitr::opts_chunk$set(dev = c("png", "pdf"), fig.path = figdir)
```


# Data Import:
Load processed Data from DrugCombi_QC.Rmd and DrugCombi_AddOmics (moelcular data on patient samples)
```{r}
load(file.path(outdir, paste0("MarinaCombiDataAfterQC_",today,".RData")))
load(file.path(outdir, paste0("OmicsData",today,".RData")))
```

```{r}
# discard outliers for further analyses, only use averaged values for replicates
df4ana <- df4anaAvreplicatesRmOutliers
rm(df4anaAvreplicatesRmOutliers, df4anaAvreplicates)
```

# Introduction
To explore the synergistic effect of two drug we use the following measures:

- Additive effect model: Compare the measured viability $v_{BC}$ to $v^*_{BC} = v_B v_c$ and use $SI = $v^*_{BC}- v_{BC}$ ( this is larger than 0 if the viability vlaues of the combinatino are lower than the additive model (-> Synergy)), summarized across individual concentrations by AUC
- Bliss Index: $\frac{e_B + e_C -e_B e_C}{e_{BC}}$, summarized across individual concentrations by averaging
- Strongest singel agent Index: $\frac{max(e_B, e_C)}{e_{BC}}$, summarized across individual concentrations by averaging

Denoting the viability values as $v_X$ for base drug alone (B), combination compund alone (C) and the combination (BC) and $e_X = 1- max(1, v_X))$.

# Add combination indices for individual measurements
```{r}
# Bliss and SI are based on identical predicitions for combi, difference only  v* - v instea of (1 - v*)/(1-v) and no trhesolding at 1.

# dfsyn %<>% dplyr::rename(viabB = effectB,
#                   viabC = effectC,
#                   viabBC = effectBC)
# dfsyn %<>% mutate(effB = 1- pmin(viabB,1),
#                   effC = 1- pmin(viabC,1),
#                   effBC = 1- pmin(viabBC, 1))
# dfsyn %<>% mutate(viabBC_add = pmin(1,viabB) * pmin(1,viabC),
#                   effBC_add = 1- viabBC_add)
# dfsyn %<>% mutate(addModelSI =  effBC_add/effBC,
#                   BlissCI = (effB + effC - effB * effC)/effBC) 
# 
# dfsyn %<>% mutate(hsaCI = ifelse(effBC!=0, pmax(effB, effC)/effBC,1))
# dfsyn %<>% mutate(BlissCI = ifelse(effBC!=0,(effB + effC - effB * effC)/effBC, 1))
```

```{r CI}
dfsyn <- df4ana
rm(df4ana)
# rename effect to viab and add 'effect' as defined above (1-viability)
dfsyn %<>% dplyr::rename(viabB = effectB,
                  viabC = effectC,
                  viabBC = effectBC)
dfsyn %<>% mutate(effB = 1- pmin(viabB,1),
                  effC = 1- pmin(viabC,1),
                  effBC = 1- pmin(viabBC, 1))

# add theoretical additive effect of drug combinations
dfsyn %<>% mutate(viabBC_add = viabB * viabC)

# add combination synergy measures
#FIX ME: DEVISION BY VERY SMALL EFFECTS LEADS TO VERY LARGE VALUES....
dfsyn %<>% mutate(addModelSI =  viabBC_add- viabBC) 
dfsyn %<>% mutate(hsaCI = ifelse(effBC!=0, pmax(effB, effC)/effBC,1))
dfsyn %<>% mutate(BlissCI = ifelse(effBC!=0,(effB + effC - effB * effC)/effBC, 1))

# viz
hist(log(dfsyn$hsaCI))
hist(log(dfsyn$BlissCI))
hist(dfsyn$addModelSI)
```

# Summarize concentrations
If we use the area between the measured and additive effect curve as aggregated SI caluclated byb a simple Trapez rule and defining the concentrations to be equisidstant the highest and lowest concentrations will get half the weight compared to the others ($AUC = 0.5v_{c1} +  v_{c2} +  v_{c3} +  v_{c4} + 0.5v_{c5}$ ).
If we use the acutal concentration values the measure is influenced by the absolute range of concentrations used for a compound and smaller concentration values that are typically closer together alos get smaller weights.
--> Use a simple average across concentrations giving all concentrations same influence on the snergy index.

```{r}
# by mean across concentrations
dfsynSummaryConc <- group_by(dfsyn, CDrugAbrv, CDrugName, CDrugID, CDrugConc, CDrugNameLong,
                 BDrugName, BDrugID, PatientID) %>%
  summarize(BlissCImean = mean(BlissCI, na.rm =TRUE),
                    hsaCImean = mean(hsaCI, na.rm =TRUE),
                    addModelSImean = mean(addModelSI, na.rm =TRUE),
                    hsaCIse = sd(BlissCI, na.rm =TRUE)/ sqrt(sum(!is.na(BlissCI))),
                    BlissCIse = sd(hsaCI, na.rm =TRUE)/ sqrt(sum(!is.na(hsaCI))),
                    addModelSIse = sd(addModelSI, na.rm =TRUE)/ sqrt(sum(!is.na(addModelSI)))) %>%
  ungroup()
  
# by median across patients
dfsynSummaryConcPat <- dfsynSummaryConc %>% 
  group_by(CDrugAbrv, CDrugName, CDrugID, CDrugConc, CDrugNameLong, BDrugName, BDrugID) %>%
  summarize(BlissCImed = median(BlissCImean),
                    hsaCImed = median(hsaCImean),
                    addModelSImed = median(addModelSImean)) %>%
  ungroup()

# by median across patients
dfsynSummaryPat <- dfsyn %>% 
  group_by(CDrugAbrv, CDrugName, CDrugID, CDrugConc, CDrugNameLong, BDrugName, BDrugID, BDrugConc, BDrugConcId) %>%
  summarize(BlissCImed = median(BlissCI, na.rm = TRUE),
                    hsaCImed = median(hsaCI, na.rm =TRUE),
                    addModelSImed = median(addModelSI, na.rm =TRUE)) %>%
  ungroup()
```

# Heatmap SI drug-drug combination
```{r SI_heatmap, fig.height=9}
SImat <- dfsynSummaryConcPat %>%
  select(CDrugAbrv,BDrugName, addModelSImed) %>%
  dplyr::rename(SI = addModelSImed) %>% spread(key=BDrugName, value=SI) %>% 
  column_to_rownames("CDrugAbrv") %>% as.matrix()

myHeatmap(t(SImat), colors=c("navy", "lightblue", "white", "orangered","darkred"))
colSums(SImat>0)

pheatmap(t(SImat), breaks = seq(0, max(SImat), length.out = 100), main = "SI",
         color = colorRampPalette(rev(brewer.pal(n = 7, name ="RdYlBu")))(100))

dfsynSummaryConcPat %>%
  select(CDrugAbrv,BDrugName, addModelSImed) %>% filter(CDrugAbrv !="Duvelisib") %>%
  rename(SI = addModelSImed) %>% arrange(-SI)
```

# Heatmap HSA
```{r CIhsa_heatmap, fig.height=9}
SImat <- dfsynSummaryConcPat %>%
  select(CDrugAbrv,BDrugName, hsaCImed) %>%
  dplyr::rename(SI = hsaCImed) %>% spread(key=BDrugName, value=SI) %>% 
  column_to_rownames("CDrugAbrv") %>% as.matrix()

pheatmap(t(SImat), breaks = seq(0,1,0.01), main = "Highest single agent Index",
                  color = colorRampPalette(brewer.pal(n = 7, name ="RdYlBu"))(100))
```

# Heatmap Bliss
Values below 1 indicate synergy.
```{r CIBliss_heatmap, fig.height=9}
SImat <- dfsynSummaryConcPat %>%
  select(CDrugAbrv,BDrugName, BlissCImed) %>%
  dplyr::rename(SI = BlissCImed) %>% spread(key=BDrugName, value=SI) %>% 
  column_to_rownames("CDrugAbrv") %>% as.matrix()

pheatmap(t(SImat), breaks = seq(0,1,0.01), main = "Bliss Index",
         color = colorRampPalette(brewer.pal(n = 7, name ="RdYlBu"))(100))
```


<!-- # Linear model with interaction terms -->
<!-- Only include combination with at least 10 samples -->
<!-- ```{r} -->
<!-- # fit a linear model for each drug-drug combi and concentration (conc are not independent) across samples -->
<!-- dfn <- dfsyn %>% group_by(CDrugAbrv, BDrugName, BDrugConcId) %>% summarise(nsamples = n()) %>% filter(nsamples>=10) -->

<!-- dflm <- dfsyn %>% left_join(dfn, by= c("CDrugAbrv", "BDrugName", "BDrugConcId")) %>% -->
<!--   filter(nsamples >= 10) %>% -->
<!--   group_by(CDrugAbrv, BDrugName, BDrugConcId) %>% -->
<!--   do(lm.out = lm(log(.$viabBC) ~ log(.$viabB) + log(.$viabC) + log(.$viabC)*log(.$viabB))) -->
<!-- dflm %<>% mutate(intercept = lm.out$coefficient[1], -->
<!--                  betaB = lm.out$coefficient[2], -->
<!--                  betaC = lm.out$coefficient[3], -->
<!--                  betaBC = lm.out$coefficient[4], -->
<!--                  pBC = coefficients(summary(lm.out))[4,4]) -->
<!-- ``` -->

<!-- ```{rlm_heatmap, fig.height=9} -->
<!-- betamat <- dflm %>% group_by(CDrugAbrv, BDrugName) %>% -->
<!--   summarize(cidx = which.max(betaBC), BDrugConcId = BDrugConcId[cidx],pBC = pBC[cidx], betaBC = betaBC[cidx]) %>% -->
<!--   select(CDrugAbrv,BDrugName, betaBC) %>% -->
<!--   spread(key=BDrugName, value=betaBC) %>%  -->
<!--   column_to_rownames("CDrugAbrv") %>% as.matrix() -->

<!-- myHeatmap(t(betamat), colors=c("navy", "blue", "white", "red","darkred")) -->

<!-- pmat <- dflm %>% group_by(CDrugAbrv, BDrugName) %>% -->
<!--   summarize(cidx = which.max(betaBC), BDrugConcId = BDrugConcId[cidx],pBC = pBC[cidx], betaBC = betaBC[cidx]) %>% -->
<!--   select(CDrugAbrv,BDrugName, pBC) %>% -->
<!--   spread(key=BDrugName, value=pBC) %>%  -->
<!--   column_to_rownames("CDrugAbrv") %>% as.matrix() -->

<!-- myHeatmap(-log10(t(pmat)), colors=c("white", "white", "white", "grey","black")) -->

<!-- dflm$padj = p.adjust(dflm$pBC, method = "BH") -->
<!-- filter(dflm, padj < 0.05) -->
<!-- ``` -->

# Volanoe SI vs TP53 and Ibrutinib

```{r}
alpha <- 0.05
minOcc <- 3
```

```{r test_SI}
dfSI_gen_test <- dfsynSummaryConc %>%
  select(BDrugName, CDrugAbrv, addModelSImean, PatientID) %>%
  left_join(gather(dfMuts4testing, key="mutation", value="status", -PatientID), by="PatientID")
dfSI_gen_test %<>% mutate(combi =paste(CDrugAbrv, BDrugName, sep="+"))

# only inclduing durg-drug combinations with at least 3 samples from mutated and unmutated group
df_pvals_SI_vs_muts <- group_by(dfSI_gen_test, CDrugAbrv, BDrugName, mutation, combi) %>%
  summarize(n_muts= sum(status==1), n_wt =sum(status==0),
            pval = ifelse(n_muts<minOcc | n_wt<minOcc, NA,t.test(addModelSImean ~ status, var.equal= TRUE)$p.value),
            mean.diff = ifelse(n_muts<minOcc | n_wt<minOcc, NA,diff(t.test(addModelSImean ~ status, var.equal= TRUE)$estimate))) %>%
  filter(!is.na(pval)) %>% ungroup()
df_pvals_SI_vs_muts$pval.adj <-  p.adjust(df_pvals_SI_vs_muts$pval, method="BH")
hist(df_pvals_SI_vs_muts$pval)
filter(df_pvals_SI_vs_muts, pval.adj < alpha)
```

```{r volcanoe_SI}
ggplot(df_pvals_SI_vs_muts, aes(x = mean.diff, y = -log10(pval.adj), col = mean.diff > 0)) +
  geom_point(alpha=0.7) +
  ggrepel::geom_label_repel(aes(label=ifelse(pval.adj < alpha, combi, "")), label.size= 0, min.segment.length = 0, fill = NA) +
  scale_color_manual(values= c("deeppink", "navy")) + guides(col=FALSE) +
  theme_bw() + geom_hline(yintercept = -log10(alpha), lty= "dashed", alpha=0.3) +
  xlab("Mean SI difference")  +facet_wrap(~mutation)
```

# Ibrutinib (100nM) + Navitoclax

## Response curves
Viability response curves (mean across patients) for single drugs, combination and additive effect model.
```{r Ibrutinib (100nM)_Navitoclax_Curves, fig.width=9}
plotResponseCurves(df =dfsyn, drC = "Ibrutinib (100nM)", drB = "Navitoclax", annoSI = TRUE)
```

## Boxplots CI
```{r Ibrutinib (100nM)_Navitoclax_CIboxplots_Bliss, fig.width=4}
plotBoxplotCI(df =dfsyn, drC = "Ibrutinib (100nM)", drB = "Navitoclax", CI_type = "Bliss")
```

```{r Ibrutinib (100nM)_Navitoclax_CIboxplots_HSA, fig.width=4}
plotBoxplotCI(df =dfsyn, drC = "Ibrutinib (100nM)", drB = "Navitoclax", CI_type = "hsa")
```

```{r Ibrutinib (100nM)_Navitoclax_CIboxplots_SI, fig.width=4}
plotBoxplotCI(df =dfsyn, drC = "Ibrutinib (100nM)", drB = "Navitoclax", CI_type = "SI")
```
## Waterfall CI

```{r Ibrutinib (100nM)_Navitoclax_CIwaterfall, fig.width=10, fig.height=3}
plotWaterfallCI(dfsynSummaryConc, drC = "Ibrutinib (100nM)", drB = "Navitoclax", CI_type = "SI")
```

## Scatter

```{r Ibrutinib (100nM)_Navitoclax_Scatter, fig.width=9}
plotScattter(dfsyn,  drC = "Ibrutinib (100nM)", drB = "Navitoclax")
```


# Ibrutinib (100nM) + Afatinib

## Response curves
Viability response curves (mean across patients) for single drugs, combination and additive effect model.
```{r Ibrutinib (100nM)_Afatinib_Curves, fig.width = 12}
plotResponseCurves(df =dfsyn, drC = "Ibrutinib (100nM)", drB = "Afatinib", annoSI = TRUE)
```

## Boxplots CI
```{r Ibrutinib (100nM)_Afatinib_CIboxplots_Bliss, fig.width=4}
plotBoxplotCI(df =dfsyn, drC = "Ibrutinib (100nM)", drB = "Afatinib", CI_type = "Bliss")
```

```{r Ibrutinib (100nM)_Afatinib_CIboxplots_HSA, fig.width=4}
plotBoxplotCI(df =dfsyn, drC = "Ibrutinib (100nM)", drB = "Afatinib", CI_type = "hsa")
```

```{r Ibrutinib (100nM)_Afatinib_CIboxplots_SI, fig.width=4}
plotBoxplotCI(df =dfsyn, drC = "Ibrutinib (100nM)", drB = "Afatinib", CI_type = "SI")
```

## Waterfall CI

```{r Ibrutinib (100nM)_Afatinib_CIwaterfall, fig.width=10, fig.height=3}
plotWaterfallCI(dfsynSummaryConc, drC = "Ibrutinib (100nM)", drB = "Afatinib", CI_type = "SI")
```

## Scatter
```{r Ibrutinib (100nM)_Afatinib_Scatter, fig.width=9}
plotScattter(dfsyn,  drC = "Ibrutinib (100nM)", drB = "Afatinib")
```


# Ibrutinib (100nM) + Venetoclax

## Response curves
Viability response curves (mean across patients) for single drugs, combination and additive effect model.
```{r Ibrutinib (100nM)_Venetoclax_Curves, fig.width=12}
plotResponseCurves(df =dfsyn, drC = "Ibrutinib (100nM)", drB = "Venetoclax", annoSI = TRUE)
```

## Boxplots CI
```{r Ibrutinib (100nM)_Venetoclax_CIboxplots_Bliss, fig.width=4}
plotBoxplotCI(df =dfsyn, drC = "Ibrutinib (100nM)", drB = "Venetoclax", CI_type = "Bliss")
```

```{r Ibrutinib (100nM)_Venetoclax_CIboxplots_HSA, fig.width=4}
plotBoxplotCI(df =dfsyn, drC = "Ibrutinib (100nM)", drB = "Venetoclax", CI_type = "hsa")
```

```{r Ibrutinib (100nM)_Venetoclax_CIboxplots_SI, fig.width=4}
plotBoxplotCI(df =dfsyn, drC = "Ibrutinib (100nM)", drB = "Venetoclax", CI_type = "SI")
```

## Waterfall CI
```{r Ibrutinib (100nM)_Venetoclax_CIwaterfall, fig.width=10, fig.height=3}
plotWaterfallCI(dfsynSummaryConc, drC = "Ibrutinib (100nM)", drB = "Venetoclax", CI_type = "SI")
```

## Scatter
```{r Ibrutinib (100nM)_Venetoclax_Scatter, fig.width=9}
plotScattter(dfsyn,  drC = "Ibrutinib (100nM)", drB = "Venetoclax")
```

# Ibrutinib (100nM) + Fludarabine

## Response curves
Viability response curves (mean across patients) for single drugs, combination and additive effect model.
```{r Ibrutinib (100nM)_Fludarabine_Curves, fig.width=12, fig.height=12}
plotResponseCurves(df =dfsyn, drC = "Ibrutinib (100nM)", drB = "Fludarabine", annoSI = TRUE)
```

## Boxplots CI
```{r Ibrutinib (100nM)_Fludarabine_CIboxplots_Bliss, fig.width=4}
plotBoxplotCI(df =dfsyn, drC = "Ibrutinib (100nM)", drB = "Fludarabine", CI_type = "Bliss")
```

```{r Ibrutinib (100nM)_Fludarabine_CIboxplots_HSA, fig.width=4}
plotBoxplotCI(df =dfsyn, drC = "Ibrutinib (100nM)", drB = "Fludarabine", CI_type = "hsa")
```

```{r Ibrutinib (100nM)_Fludarabine_CIboxplots_SI, fig.width=4}
plotBoxplotCI(df =dfsyn, drC = "Ibrutinib (100nM)", drB = "Fludarabine", CI_type = "SI")
```

## Waterfall CI
```{r Ibrutinib (100nM)_Fludarabine_CIwaterfall, fig.width=10, fig.height=3}
plotWaterfallCI(dfsynSummaryConc, drC = "Ibrutinib (100nM)", drB = "Fludarabine",
                CI_type = "SI", annotate = c("TP53", "IGHV"))
```

## Scatter
```{r Ibrutinib (100nM)_Fludarabine_Scatter, fig.width=9}
plotScattter(dfsyn,  drC = "Ibrutinib (100nM)", drB = "Fludarabine")
```

## SI boxplot by TP53 and IGHV
```{r Ibrutinib (100nM)_Fludarabine_SI_byMuts}
df4plot <- dfsynSummaryConc %>% filter(CDrugAbrv == "Ibrutinib (100nM)",
                                       BDrugName == "Fludarabine") 
# myComp <- lapply(1:6, function(x) combn(unique(df4plot$status),2)[,x])

df4plot %<>% left_join(dfMuts4testing)
df4plot %<>% mutate(IGHV = ifelse(IGHV == 0, "U-CLL", "M-CLL"))
df4plot %<>% mutate(TP53 = ifelse(TP53 == 0, "TP53-wt", "TP53-mut"))
df4plot %<>% mutate(status = paste(IGHV, TP53, sep = ", "))
ggplot(df4plot, aes(x=status, y=addModelSImean)) +
    ggbeeswarm::geom_beeswarm(aes(col = IGHV, shape=TP53), alpha =0.7) + 
    geom_boxplot(outlier.shape = NA, alpha=0.2, width = 0.3) +
    geom_hline(yintercept = 0, lty = "dashed") +
    theme_bw()  + xlab("Genetic group") +
    ylab("Synergy index (SI)") +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank())   +
  ggpubr::stat_compare_means(aes(group = status))
```


# Ibrutinib + MK-2206
## Response curves
```{r Ibrutinib (100nM)_MK-2206_Curves, fig.width=12}
plotResponseCurves(df =dfsyn, drC = "Ibrutinib (100nM)", drB = "MK-2206", annoSI = TRUE)
```
## Boxplots CI
```{r Ibrutinib (100nM)_MK-2206_CIboxplots_Bliss, fig.width=4}
plotBoxplotCI(df =dfsyn, drC = "Ibrutinib (100nM)", drB = "MK-2206", CI_type = "Bliss")
```

```{r Ibrutinib (100nM)_MK-2206_CIboxplots_HSA, fig.width=4}
plotBoxplotCI(df =dfsyn, drC = "Ibrutinib (100nM)", drB = "MK-2206", CI_type = "hsa")
```


# Ibrutinib + Everolimus
## Response curves
```{r Ibrutinib (100nM)_Everolimus_Curves, fig.width=12}
plotResponseCurves(df =dfsyn, drC = "Ibrutinib (100nM)", drB = "Everolimus", annoSI = TRUE)
```
## Boxplots CI
```{r Ibrutinib (100nM)_Everolimus_CIboxplots_Bliss, fig.width=4}
plotBoxplotCI(df =dfsyn, drC = "Ibrutinib (100nM)", drB = "Everolimus", CI_type = "Bliss")
```

```{r Ibrutinib (100nM)_Everolimus_CIboxplots_HSA, fig.width=4}
plotBoxplotCI(df =dfsyn, drC = "Ibrutinib (100nM)", drB = "Everolimus", CI_type = "hsa")
```

# Ibrutinib + Nutlin-3
## Response curves
```{r Ibrutinib (100nM)_Nutlin-3_Curves, fig.width=9}
plotResponseCurves(df =dfsyn, drC = "Ibrutinib (100nM)", drB = "Nutlin-3", annoSI = TRUE)
```
## Boxplots CI
```{r Ibrutinib (100nM)_Nutlin-3_CIboxplots_Bliss, fig.width=4}
plotBoxplotCI(df =dfsyn, drC = "Ibrutinib (100nM)", drB = "Nutlin-3", CI_type = "Bliss")
```

```{r Ibrutinib (100nM)_Nutlin-3_CIboxplots_HSA, fig.width=4}
plotBoxplotCI(df =dfsyn, drC = "Ibrutinib (100nM)", drB = "Nutlin-3", CI_type = "hsa")
```


# Overview BCRi
```{r}
BCRi <- filter(DrugMetaList$DrugInfoCombi, pathway == "B-cell receptor")
BCRi <- BCRi$niceName
# only use one concentration of Ibrutinib (Ibrutinib (100nM) has more samples)
BCRi <- BCRi[BCRi != "Ibrutinib (10nM)"]
```

## Waterfall plot SI
```{r BCRi_WaterfallSI, fig.height= 9}
B_BCRi <- DrugMetaList$DrugInfoBase %>% filter(grepl("Btk", Target) | grepl("PI3Kδ", Target) | grepl("Syk", Target))
B_BCL2 <-  DrugMetaList$DrugInfoBase %>% filter(grepl("Bcl-2", Target))

df_orderedbyIbru100 <-  filter(dfsynSummaryConcPat, CDrugAbrv == "Ibrutinib (100nM)") %>% arrange(addModelSImed)
dfsynSummaryConcPat$BDrugName <- factor(dfsynSummaryConcPat$BDrugName, df_orderedbyIbru100$BDrugName)


dfsynSummaryConcPat %>% filter(CDrugAbrv %in% BCRi) %>%
  mutate(type = ifelse(BDrugName %in% B_BCRi$niceName, "BCRi", 
                         ifelse(BDrugName %in% B_BCL2$niceName, "BCL2i", "other"))) %>%
  mutate(CDrugAbrv = factor(CDrugAbrv, levels = c("Ibrutinib (100nM)", BCRi[BCRi!="Ibrutinib (100nM)"]))) %>%
  ggplot(aes(x = BDrugName, y = addModelSImed, fill=type)) + geom_bar(stat="identity") +
  facet_wrap(~CDrugAbrv, ncol=1) +
  theme(axis.text.x = element_text(angle=60, vjust=1, hjust=1),
        axis.ticks.x = element_blank(),
        panel.background = element_rect(fill = "white"))  +
  ylab("Synergy Index (SI)") + xlab("Base compound") 
```

## Response curves
```{r BCRi_ResponseCurves, fig.height= 9}
plotMultipleBResponseCurves(dfsyn,  drC = "Ibrutinib (100nM)", drsB = c("Spebrutinib", "Duvelisib", "R406", "AZD7762", "Dasatinib", "PF 477736", "Idelalisib"), annoSI = TRUE)
```

##Scatterplot
```{r BCRi_Scatterplot, fig.width=9}
plotScattterVsCombi(dfsyn,  drC = "Ibrutinib (100nM)", drB = "Idelalisib")
```

## Boxplots CI
```{r Ibrutinib (100nM)_Idelalisib_CIboxplots_Bliss, fig.width=4}
plotBoxplotCI(df =dfsyn, drC = "Ibrutinib (100nM)", drB = "Idelalisib", CI_type = "Bliss")
```


# Counterscreen Afatinib
## Waterfall SI
```{r AfatinibCounter_WaterfallSI, fig.height= 10, fig.height=3}
B_BCRi <- DrugMetaList$DrugInfoBase %>% filter(grepl("Btk", Target) | grepl("PI3Kδ", Target) | grepl("Syk", Target))

df_orderedbyAfatinib <-  filter(dfsynSummaryConcPat, CDrugAbrv == "Afatinib") %>% arrange(addModelSImed)
dfsynSummaryConcPat$BDrugName <- factor(dfsynSummaryConcPat$BDrugName, df_orderedbyAfatinib$BDrugName)


dfsynSummaryConcPat %>% filter(CDrugAbrv == "Afatinib") %>%
  mutate(type = ifelse(BDrugName %in% B_BCRi$niceName, "BCRi", "other")) %>%
  ggplot(aes(x = BDrugName, y = addModelSImed, fill=type)) + geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle=60, vjust=1, hjust=1),
        axis.ticks.x = element_blank(),
        panel.background = element_rect(fill = "white"))  +
  ylab("Synergy Index (SI)") + xlab("Base compound")  +guides(fill=FALSE) +
  scale_fill_manual(values = c(BCRi = "forestgreen", other= "gray"))
```

## Afatinib + Ibrutinib (counterscreen)

### Scatterplot
```{r Afatinib_Ibrutinib_Scatterplot, fig.width=9}
plotScattterVsCombi(dfsyn,  drC = "Afatinib", drB = "Ibrutinib")
```

### Waterfall CI
```{r Afatinib_Ibrutinib_CIwaterfall, fig.width=10, fig.height=3}
plotWaterfallCI(dfsynSummaryConc, drC = "Afatinib", drB = "Ibrutinib", CI_type = "SI")
```


# Data export
```{r}
save(dfsyn, dfsynSummaryConc, dfsynSummaryConcPat, dfsynSummaryPat, file=paste0("out/MarinaCombiSyns_",today,".RData"))
```


```{r}
sessionInfo()
```


