---
title: "CLL drug combinations ex-vivo: Quality Control"
author: "Britta Velten"
date: "30 August 2018"
output:
  BiocStyle::html_document:
    toc: true
    toc_depth: 3
params: 
    today: 180914
---

#Introduction

Takes the data objects created in `DrugCombi_DataImport.Rmd` and performs further QC on the data.

```{r, echo=F}
library(knitr )
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
options(stringAsFactors = FALSE)
```

```{r, echo=F}
library(ggplot2)
library(RColorBrewer)
library(Biobase)
library(abind)
library(grid)
library(gtable)
library(reshape2)
library(gridExtra)
require(pracma)
library(reshape2)
library(dplyr)
library(pheatmap)
library(magrittr)
library(lattice)
library(tidyverse)
```

```{r}
datadir <- "/Volumes/huber/projects/nct/cll/RawData/DrugScreens/Marina_CombiScreen/"
outdir = "out"
today <- params$today
figdir = paste0("figs", today, "/figuresQC/")
if(!dir.exists(figdir)) dir.create(figdir)
knitr::opts_chunk$set(dev = c("png", "pdf"), fig.path = figdir)
```


# Preparations
# Data Import 
Load processed Data from `DrugCombi_DataImport.Rmd`.
```{r}
load(file.path(outdir, paste0("MarinaCombiData_",today,".RData")))
```

Loaded dataframe from ImportedData:

* `CompleteDF` contains the raw and normalized values per well/plate/screen
* `df4ana` contains the normalized values for each single drug and drug-drug combination per sample and concentration)
* `DrugMetaList` contains info about the drugs used as base and combi compounds
* `SetUpList` contains the plate setups for the screens

## Setting colors for patients.
```{r}
colset = grDevices::colors()[!grepl("^gr(a|e)y", grDevices::colors())]
patcol = unique(df4ana$PatientID)
set.seed(2808)
patcol = setNames(colset[sample.int(length(colset),length(patcol))], nm=patcol)
```

Color legend for patient:
```{r legendPatients, eval=F, echo=F}
ggplot(data=data.frame(X=names(patcol), Y=1)) +
  geom_bar(aes(x=X, y=Y, fill=X), stat="identity") +
  scale_fill_manual(values=patcol) + guides(fill=FALSE) +
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.5)) +
  xlab("") + ylab("")
```



# Plate plots
Produces plot of the raw and normalized viability values for each plate based on the data frame containting the per well values (`CompleteDF`).

Raw values:
```{r plateplots_raw, fig.width=20, fig.height=100, eval=FALSE}
ggplot(CompleteDF, aes(y=factor(WellLetter, levels=letters[16:1]), x=WellNo)) +
  facet_wrap(~ PlateID, ncol=4) + geom_tile(aes(fill=rawValue)) +
  scale_fill_gradient(low="white", high="black")
```

Normalized values:
```{r plateplots_norm, fig.width=20, fig.height=100, eval=FALSE}
ggplot(CompleteDF, aes(y=factor(WellLetter, levels=letters[16:1]), x=WellNo)) +
  facet_wrap(~ PlateID, ncol=4) + geom_tile(aes(fill=normalizedValue)) +
  scale_fill_gradient(low="white", high="black")
```


# Check reproducibility between replicates and average
Until now the dataframe contains two replicates for each concentration and drug combination in screen 1 and 2. Here, we check the agreement between replicates and then continue with averages.

Replicates are present in screen 1 and 2: Here every measurement has one replicate
```{r}
for (screen in sort(as.character(unique(df4ana$ScreenNo)))){
df <- filter(df4ana, ScreenNo==screen)  
checkDup <- select(df, PatientID, CDrugID, BDrugID, ScreenNo, CDrugConc, BDrugConcId)
dupIDX <- which(duplicated(checkDup)|duplicated(checkDup, fromLast=T))
print(paste(screen,": ", length(dupIDX), " out of ", nrow(df), " duplicated", sep=""))
}
```

Annotate replicates
```{r}
df4ana$replicate <- sapply(duplicated(select(df4ana, PatientID, CDrugID,
                                             BDrugID, ScreenNo, CDrugConc, BDrugConcId)),
                           function(bool) ifelse(bool,2,1))
sum(df4ana$replicate==2)
df4ana %<>% mutate(patrep = paste(PatientID, replicate, sep="_"))

CompleteDF$replicate  <- sapply(duplicated(select(CompleteDF, PatientID, BaseDrugID,
                                                  BaseDrugConcID, CombiDrug, ScreenDate,
                                                  BaseDrugConc, BaseDrugName, ScreenNo,
                                                  CombiDrugConc, CombiDrugID,CombiDrug_longname)),
                                function(bool) ifelse(bool,2,1))
sum(CompleteDF$replicate==2)
CompleteDF %<>% mutate(patrep = paste(PatientID, replicate, sep="_"))
```

```{r replicateConsensus_raw_all, eval=FALSE, echo=FALSE}
df_reps <- filter(CompleteDF, ScreenNo %in% c("SS1","SS2")) %>%
  filter(BaseDrugID != "DM") %>% # remove control wells (no Drug B nor C)
  filter(BaseDrugID != "DM+") %>% # remove wells with only C (muliple wells-average)
  select(-starts_with("Well")) %>%
  select(-patrep) %>%
  select(-normalizedValue) %>%
  mutate(replicate = paste0("rep",replicate)) %>%
  spread(key="replicate", value="rawValue")

ggplot(df_reps, aes(x=rep1, y=rep2, col=as.character(PatientID))) + geom_point() +
   scale_color_manual(values = patcol) +
  geom_abline(intercept=0,slope=1, lty="dashed") +
  geom_text(x =100000, y =600000, label = paste("r =", round(cor(df_reps$rep1,df_reps$rep2),2)), col="black", size =5) +
  theme_bw() + guides(col=FALSE) + xlab("raw intensities  (replicate 1)") + ylab("raw intensities (replicate 2)") +coord_fixed()
```

Check agreement between replicates for the drug-drug combination values (rawValueBC or effectBC), output: "replicateConsensus.pdf"
```{r replicateConsensus_raw}
# raw drug-drug combination intensities
df_reps <- filter(df4ana, ScreenNo %in% c("SS1","SS2")) %>%
  select(-starts_with("effect")) %>%
  select(-patrep) %>%
  mutate(replicate = paste0("rep",replicate)) %>%
  spread(key="replicate", value="rawValueBC")

ggplot(df_reps, aes(x=rep1, y=rep2, col=PatientID)) + geom_point() +
   scale_color_manual(values = patcol) +
  geom_abline(intercept=0,slope=1, lty="dashed") +
  geom_text(x =100000, y =600000,
            label = paste("r =", round(cor(df_reps$rep1,df_reps$rep2),2)),
            col="black", size =5) +
  theme_bw() + guides(col=FALSE) + xlab("raw intensities  (replicate 1)") +
  ylab("raw intensities (replicate 2)") + coord_fixed()
```

```{r replicateConsensus_normalized}
# normalized drug-drug combination viabilities
df_reps <- filter(df4ana, ScreenNo %in% c("SS1","SS2")) %>%
  select(-c(effectB, effectC, patrep, rawValueBC)) %>%
  mutate(replicate = paste0("rep",replicate)) %>%
  spread(key="replicate", value="effectBC")

ggplot(df_reps, aes(x=rep1, y=rep2, col=PatientID)) + geom_point() +
  scale_color_manual(values = patcol)+ theme_bw() + 
    geom_abline(intercept=0,slope=1, lty="dashed") +
    geom_text(x =0.3, y =1.2,
              label = paste("r =", round(cor(df_reps$rep1,df_reps$rep2),2)),
              col="black", size =5) +
  guides(col=FALSE) + xlab("normalized viabilities (replicate 1)") +
  ylab("normalized viabilities (replicate 2)")+coord_fixed()

# normalized base drug viabilities
df_reps <- filter(df4ana, ScreenNo %in% c("SS1","SS2")) %>%
  select(-c(effectBC, effectC, patrep, rawValueBC)) %>%
  mutate(replicate = paste0("rep",replicate)) %>%
  spread(key="replicate", value="effectB")

ggplot(df_reps, aes(x=rep1, y=rep2, col=PatientID)) + geom_point() +
  scale_color_manual(values = patcol)+ theme_bw() + 
    geom_abline(intercept=0,slope=1, lty="dashed") +
      geom_text(x =0.3, y =1.2,
                label = paste("r =", round(cor(df_reps$rep1,df_reps$rep2),2)),
                col="black", size =5) +
  guides(col=FALSE) + xlab("normalized viabilities (replicate 1)") +
  ylab("normalized viabilities (replicate 2)")+ coord_fixed()

# normalized combi drug viabilities
df_reps <- filter(df4ana, ScreenNo %in% c("SS1","SS2")) %>%
  select(-c(effectBC, effectB, patrep, rawValueBC)) %>%
  mutate(replicate = paste0("rep",replicate)) %>%
  spread(key="replicate", value="effectC")

ggplot(df_reps, aes(x=rep1, y=rep2, col=PatientID)) + geom_point() +
  scale_color_manual(values = patcol)+ theme_bw() + 
    geom_abline(intercept=0,slope=1, lty="dashed") +
      geom_text(x =0.7, y =1.2,
                label = paste("r =", round(cor(df_reps$rep1,df_reps$rep2),2)),
                col="black", size =5) +
  guides(col=FALSE) + xlab("normalized viabilities (replicate 1)") +
  ylab("normalized viabilities (replicate 2)")+ coord_fixed()
```

Note: The last plot should be a straight line as same DM+ wells used for effectC and only one value per plate

As screen 1 is not included for tha analysis sue to the lack of a good control plate, we limit the replicate analysis to screen 2:
Check agreement between replicates, output: "replicateConsensus.pdf"
```{r replicateConsensus_raw_screen2}
# raw drug-drug combination intensities
df_reps <- filter(df4ana, ScreenNo %in% c("SS2")) %>%
  select(-starts_with("effect")) %>%
  select(-patrep) %>%
  mutate(replicate = paste0("rep",replicate)) %>%
  spread(key="replicate", value="rawValueBC") 

ggplot(df_reps, aes(x=rep1, y=rep2, col=PatientID)) + geom_point(size=3) +
   scale_color_manual(values = patcol) +
  geom_abline(intercept=0,slope=1, lty="dashed") +
  geom_text(x =100000, y =600000,
            label = paste("r =", round(cor(df_reps$rep1,df_reps$rep2),2)), col="black", size =7) +
  theme_bw(base_size = 20) + guides(col=FALSE) + xlab("raw intensities  (replicate 1)") +
  ylab("raw intensities (replicate 2)") +coord_fixed()
```

Calculate average value for replicates
```{r}
df4anaAvreplicates <- df4ana %>%
  group_by(CDrugAbrv, CDrugName, CDrugID, CDrugConc, CDrugNameLong, BDrugName, BDrugID, BDrugConc, BDrugConcId, PatientID, ScreenDate,ScreenNo) %>%
  summarise(effectBC = mean(effectBC),
            effectB = mean(effectB),
            effectC = mean(effectC),
            rawValueBC = mean(rawValueBC)) %>%
  ungroup()
```

# Medpolish: Comparison of measured single effects with decomposition of combination effect by medpolish
As QC to see systematic trend in difference of single effect and combination.
Use the measured viabilities after treatment with both drugs and derive from them the single drug effect by
$\log(v_{ij})= \log(v_{i})+\log(v_{j}) + r_{ij}$ using medpolish. 
```{r medpolish}
listCombiEffectMatricesPerPatrep<-lapply(unique(df4ana$patrep), function(pr) {
  filter(df4ana, patrep==pr) %>% 
    mutate(drBc=paste(BDrugID, BDrugConcId, sep="_")) %>%  
    melt(id.vars=c("CDrugAbrv", "drBc"), measure.vars = c("effectBC")) %>% acast(CDrugAbrv~drBc)
})
listSingleEffectBVectorsPerPatrep<-lapply(unique(df4ana$patrep), function(pr) {
  dftmp<-filter(df4ana, patrep==pr) %>% mutate(drBc=paste(BDrugID, BDrugConcId, sep="_")) %>%  melt(id.vars=c( "drBc"), measure.vars = c("effectB"))
  dftmp<-dftmp[!duplicated(dftmp),] #remove duplicated rows as for each combi drug the same
  effectBvec<-dftmp$value
  names(effectBvec)<-dftmp$drBc
  effectBvec
})
listSingleEffectCVectorsPerPatrep<-lapply(unique(df4ana$patrep), function(pr) {
   dftmp<-filter(df4ana, patrep==pr) %>% mutate(drBc=paste(BDrugID, BDrugConcId, sep="_"))%>%  melt(id.vars=c( "CDrugAbrv"), measure.vars = c("effectC")) 
  dftmp<-dftmp[!duplicated(dftmp),] #remove duplicated rows as for each base drug the same
  effectCvec<-dftmp$value
  names(effectCvec)<-dftmp$CDrugAbrv
  effectCvec
})
medpolish.out.list <- lapply(listCombiEffectMatricesPerPatrep, function(mat) medpolish(log(mat)))
effectBmedpol <- lapply(medpolish.out.list, function(med) exp(med$col))
effectCmedpol <- lapply(medpolish.out.list, function(med) exp(med$row))

par(mfrow=c(3,3))
for( i in 1:length(unique(df4ana$patrep))) {
  plot(effectBmedpol[[i]], listSingleEffectBVectorsPerPatrep[[i]][names(effectBmedpol[[i]])], xlab="medpolished", ylab="measured", main=paste("effectB"))
  lines(seq(0,1,0.1),seq(0,1,0.1))
}
for( i in 1:length(unique(df4ana$patrep))) {
  plot(effectCmedpol[[i]], listSingleEffectCVectorsPerPatrep[[i]][names(effectCmedpol[[i]])], xlab="medpolished", ylab="measured", main=paste("effectC"), xlim=c(0,1.2), ylim=c(0,1.2))
    lines(seq(0,1,0.1),seq(0,1,0.1))
}
```


# Filtering

## Drop Screen 1
As no good DMSO plate available screen1 is dropped for all subseqeunt analysis
```{r}
CompleteDF <- filter(CompleteDF, ScreenNo!="SS1")
df4anaAvreplicates <- filter(df4anaAvreplicates,  ScreenNo!="SS1" )
df4ana <- filter(df4ana, ScreenNo!="SS1")
patcol <- patcol[names(patcol)%in% df4ana$PatientID]
```

## Outliers
Detect outliers: For each drug-drug combination patients with viabilities above 1.4 are removed.
```{r outliers}
max(df4anaAvreplicates$effectBC)
max(df4anaAvreplicates$effectB)
max(df4anaAvreplicates$effectC)
filter_th <- 1.4

ggplot(df4anaAvreplicates, aes(x=effectBC)) +
  geom_histogram() + 
  geom_vline(xintercept =filter_th, col="red", lty="dashed")
ggplot(df4anaAvreplicates, aes(x=effectB)) +
  geom_histogram() + 
  geom_vline(xintercept =filter_th, col="red", lty="dashed")
ggplot(df4anaAvreplicates, aes(x=effectC)) +
  geom_histogram() + 
  geom_vline(xintercept =filter_th, col="red", lty="dashed")

OutlierPoints <- filter(df4anaAvreplicates, effectB >filter_th | effectC >filter_th | effectBC>filter_th)
df4anaAvreplicatesRmOutliers <- filter(df4anaAvreplicates,
                                       effectB <=filter_th & effectC <= filter_th & effectBC <= filter_th)
```

## Are there non-active compounds?
```{r hist_effectC}
dfC <- df4ana %>% select(CDrugAbrv, effectC, PatientID) %>% filter(!duplicated(.)) 
dfC %>% ggplot(aes(x=effectC)) +geom_histogram() + facet_wrap(~ CDrugAbrv) +
  geom_vline(xintercept = 1, col = "red") +xlab("viability of combination compound alone")

df4ana %>% group_by(CDrugAbrv) %>%
  summarize(var = var(pmin(1,effectC))) %>% arrange(var)
```

```{r hist_effects_B}
dfB <- df4ana %>% select(BDrugName, BDrugConcId, effectB, PatientID) %>%
  filter(!duplicated(.)) 
dfB %>% ggplot(aes(x=effectB, fill = BDrugConcId)) +
  geom_histogram() + facet_wrap(~ BDrugName) +
  geom_vline(xintercept = 1, col = "red") +
  xlab("viability of base compound alone")

df4ana %>% group_by(BDrugName, BDrugConcId) %>%
  summarize(var = var(pmin(1,effectB)), mean = mean(effectB)) %>% 
  arrange(var) %>% group_by(BDrugName) %>%
  summarize(max_var = max(var), min_mean = min(mean)) %>% 
  arrange(max_var)
```

## Are there compounds that show a different effect in as combination drug than as base durg
```{r comparison_drC_drB}
commonDrugs <- intersect(df4ana$CDrugName, df4ana$BDrugName)
dfcombi <- DrugMetaList$DrugInfoCombi %>%
  filter(niceName %in% commonDrugs) %>%
  select(name = niceName, conc)
dfbase <- DrugMetaList$DrugInfoBase %>%
  filter(niceName %in% commonDrugs) %>%
  select(name = niceName, starts_with("c"))
full_join(dfcombi, dfbase, by = "name") %>%
  mutate()
matchConcs <- c("c4", "c5")

df4ana %>%
  filter( (BDrugName %in% commonDrugs & BDrugName == CDrugAbrv) | CDrugAbrv == "Ibrutinib (100nM)" & BDrugName == "Ibrutinib" , BDrugConcId %in% matchConcs) %>%
  mutate(label = paste(CDrugName, CDrugConc*1000, "nM")) %>%
  ggplot(aes(x=effectB, y =effectC, col = factor(BDrugConc*1000))) +
  geom_point() +facet_wrap(~label) + geom_abline(slope=1, intercept=0) +
  guides(col = guide_legend(title = "conc (B, nM)")) + coord_equal() +
  geom_hline(yintercept = 1, lty = "dashed", col="gray") + 
  geom_vline(xintercept = 1, lty = "dashed", col="gray") 

```
For Duvelisib none of the two neighboring concentration is similiar -> exclude this from the screen. 
YM155 and Venetoclax are similar in the lower concentration. Fludarabe and afatinib have no or a pro-survival effect.

Exclude Duvelisib from futher analysis, as it seems to be inactive as combination compound and not as base compound
```{r}
df4ana %<>% filter(CDrugName != "Duvelisib")
df4anaAvreplicates %<>%  filter(CDrugName != "Duvelisib")
df4anaAvreplicatesRmOutliers %<>%  filter(CDrugName != "Duvelisib")
CompleteDF %<>%  filter(CombiDrug != "Duvelisib")
```



# Data Export

Export Data to be used in further down-stream analyses

- 'DrugMetaList': contains metadata about drug
- 'CompleteDF': contains data from all plates, wells and screens, raw values and values normalized by median of DM control wells
- 'SetUpList':  contains info about plate structure for the screens
- 'df4ana': contains values for all drug-drug pairs (effectBS) with matched one-only effect from corresponding +DMSO well (effectB) or median of DM+ wells (effectC)
- 'df4anaAvrepl': df4ana with average values for replicates
- 'df4anaAvreplRemOutl': df4anaAvrepl, filtering out values above 1.4 (filter_th)
- 'patcol': color annotation for each patient
- filter_th: values above which values are filtered out


```{r}
save(DrugMetaList,CompleteDF,df4ana,SetUpList, df4anaAvreplicates, patcol, df4anaAvreplicatesRmOutliers,patcol, filter_th,
     file=paste0(outdir,"/MarinaCombiDataAfterQC_",today,".RData"))
```

Write the dataframe used for further analysis into a csv file.
```{r}
if(!dir.exists(file.path(outdir, "tables"))) dir.create(file.path(outdir, "tables"))
write.csv(CompleteDF, file= paste0(outdir,"/tables/viability_values_perplate_",today,".csv"))
write.csv(df4anaAvreplicatesRmOutliers, file= paste0(outdir,"/tables/combination_values_",today,".csv"))
```

#SessionInfo
```{r}
sessionInfo()
```

