---
title: "CLL drug combinations ex-vivo: 10x10 screens"
author: "Britta Velten"
date: "4 September 2018"
output:
  BiocStyle::html_document:
    toc: true
    toc_depth: 3
params: 
    today: 180903
---

#Introduction

Takes raw data files for 10x10 screen.

```{r, echo=F}
library(knitr )
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
options(stringAsFactors = FALSE)
```

```{r, echo=F}
library(ggplot2)
library(RColorBrewer)
library(Biobase)
library(abind)
library(grid)
library(gtable)
library(reshape2)
library(gridExtra)
require(pracma)
library(reshape2)
library(dplyr)
library(pheatmap)
library(magrittr)
library(lattice)
library(tidyverse)
library(xlsx)
```

```{r}
# setwd("~/Documents/cll/MarinaDrugComb/DrugCombi_Code4Ms/")
# datadir <- "~/Documents/cll/MarinaDrugComb/rawData"
datadir <- "/Volumes/huber/projects/nct/cll/RawData/DrugScreens/Marina_CombiScreen/"
outdir = "out"
today <- params$today
knitr::opts_chunk$set(dev = c("png", "pdf"), fig.path = paste0("figuresQC", today, "/"))
```


```{r}
datadir10x10 <- file.path(datadir,"10x10")
```


#Plate Set-up
From Prtocol get set-up of the plate (i.e. which substance in which concentration in which well)
```{r}
FileSetUp10x10 <- read.xlsx(file.path(datadir10x10,"Protokoll 10 x 10 Validierung Drugscreen2.xlsx"),2)

#get set-up for base/test drugs
  tableIdxBase <- which(FileSetUp10x10=="A", arr.ind = T)[1,]
  SetUp <- FileSetUp10x10[tableIdxBase[1]+0:15,tableIdxBase[2]+1:24]
  namesCol <- FileSetUp10x10[tableIdxBase[1]-2,tableIdxBase[2]+1:24]
  
  colnames(SetUp) <- as.matrix(namesCol)
  rownames(SetUp) <- letters[1:16]
  
  idx <- which(SetUp!="Nonsense"|is.na(SetUp), arr.ind = T)
  conc <- t(apply(SetUp,1, names))
  conc[SetUp=="DM"|is.na(SetUp)] <- NA
  letter <- apply(SetUp,2, names)
  drugs <- as.matrix(SetUp)
  SetUpBase <- data.frame(drug=as.vector(drugs), conc=as.vector(conc), 
                        letter=as.vector(letter), idx.row=idx[,1], idx.col=idx[,2])
  SetUpBase <- filter(SetUpBase, !is.na(drug))
  
#get set-up for ibrutinib
  tableIdxIbr <- which(FileSetUp10x10=="A", arr.ind = T)[2,]
  SetUp <- FileSetUp10x10[tableIdxIbr[1]+0:15,tableIdxIbr[2]+1:24]
  conc <- as.character(FileSetUp10x10[tableIdxIbr[1]+(0:15),tableIdxIbr[2]+25])
  conc[conc=="Concentration (Synergie-Substanz)"]<-NA
  SetUp <- as.matrix(SetUp)
  rownames(SetUp) <- letters[1:16]
  colnames(SetUp) <- 1:24
  #fill matrix with Ibrutinib concentration
  for (r in 1: nrow(SetUp)){
    if(!is.na(conc[r])) SetUp[r,is.na(SetUp[r,])]<-conc[r]
  }
  SetUp[grep("Ibru", SetUp)]<-sapply(strsplit(SetUp[grep("Ibru", SetUp)], split=" "),
                                   function(l) l[[2]])
  idx<-which(SetUp!="Nonsense"|is.na(SetUp), arr.ind = T)
  letter<-apply(SetUp,2, names)
  conc<-as.matrix(SetUp)
  conc[conc=="DM"]<-NA
  drug<-as.matrix(SetUp)
  drug[grep("c", drug)]<-"Ibrutinib"
  SetUpIbr<-data.frame(conc=as.vector(conc), drug=as.vector(drug),
                       letter=as.vector(letter), idx.row=idx[,1], idx.col=idx[,2])
  SetUpIbr<-filter(SetUpIbr, !is.na(drug))
```


#Read in 10x10 plates
```{r}
import10x10Data<-function(filename, SetUpBase, SetUpIbr){
  print(filename)
  data<-read.xlsx2(file.path(datadir10x10,filename),1, header = F, startRow=3)
  #info from filename
  NameComp<-strsplit(filename, "_")[[1]]
  PatientID <- NameComp[grep("P0",NameComp)] 
  TestDrugA <- NameComp[3] 
  TestDrugB <- sub(".xlsx","", NameComp[4])
  
 
  #table from excel file (ignore meta data about screen)
  tableIdx<-which(data=="<>", arr.ind = T)
  stopifnot(nrow(tableIdx)==1)
  PlateValues<-data[tableIdx[1]+1:11,tableIdx[2]+1:24]
  colnames(PlateValues)<-1:24
  rownames(PlateValues)<-letters[3:13]
  
  #data.frame match with corresponding plate set up
  value<-as.numeric(as.vector(as.matrix(PlateValues)))

  WellNo<-SetUpBase$idx.col
  stopifnot(SetUpBase$idx.col==SetUpIbr$idx.col)
  WellLetter<-SetUpBase$letter
  stopifnot(SetUpBase$letter==SetUpIbr$letter)
  #which test drug 1 is first of file name
  BaseDrug<-ifelse(SetUpBase$drug=="DM", "DM", ifelse(SetUpBase$drug==1,TestDrugA, TestDrugB))
  concBaseDrug<-SetUpBase$conc
  CombiDrug<-as.character(SetUpIbr$drug)
  concCombiDrug<-SetUpIbr$conc
  
  #pure DMSO for normalization
  pureDMSOMedian<-median(value[BaseDrug=="DM"&CombiDrug=="DM"])
  pureDMSOMean<-mean(value[BaseDrug=="DM"&CombiDrug=="DM"])
  normalizedValueDMSOControl<-value/pureDMSOMedian
  
  #get used concentrations (bug in reaad.xlsx but dilution factor is always 2 and c10 is zero, replace by hand)
  conctableIdx<-which(data=="µM"| data=="nM", arr.ind = T)
 
  concArray<-sapply(1:3, function(i) {
    concvec<-rep(0,10)
    names(concvec)<-paste("c", 1:10, sep="")
    concvec[1]<-as.numeric(as.character(data[conctableIdx[i,1]+0,conctableIdx[i,2]+1]))
    if(data[conctableIdx[i,1],conctableIdx[i,2]]=="nM")   concvec[1]<-  concvec[1]/1000
    for(j in 2:9) concvec[j]<-concvec[j-1]/2
    concvec
  })
  colnames(concArray)<-data[conctableIdx-c(1,1)]
  concArray<-cbind(concArray,DM=rep(0,10))
  colnames(concArray)[colnames(concArray)=="CAL-101"]<-"CAL101"
  colnames(concArray)[colnames(concArray)=="Abt-199"]<-"Abt199"
    colnames(concArray)[colnames(concArray)=="ABT-263"]<-"ABT263"

  print(colnames(concArray))
  
  df<-data.frame(CombiDrug=CombiDrug,
           BaseDrug=BaseDrug,
           concBaseDrug=concBaseDrug,
           concCombiDrug=concCombiDrug,
           rawValue=value, 
           pureDMSOMedian=rep(pureDMSOMedian, length(value)),
           pureDMSOMean=rep(pureDMSOMean, length(value)),
           normalizedValueDMSOControl=normalizedValueDMSOControl,
           PatientID=rep(PatientID, length(value)),
           WellNo=WellNo,
           WellLetter=WellLetter
           )
  for(i in 1:nrow(df)){
    df$concBvalue[i]<-ifelse(!is.na(df$concBaseDrug[i])&!is.na(df$BaseDrug[i]),concArray[as.character(df$concBaseDrug)[i],df$BaseDrug[i]],NA)
        df$concCvalue[i]<-ifelse(!is.na(df$concCombiDrug[i])&!is.na(df$CombiDrug[i]),concArray[as.character(df$concCombiDrug)[i],df$CombiDrug[i]],NA)
  }
  return(df)
  }
```

Different from Marina normalized values by median instead of mean (as done in other combi screens).
```{r}
allFiles<-setdiff(list.files(file.path(datadir10x10), "*.xlsx"),list.files(file.path(datadir10x10),"Protokoll"))
listOfTables<-lapply(allFiles, function(file) import10x10Data(file, SetUpBase, SetUpIbr))
for(l in 1:length(listOfTables)) listOfTables[[l]]$PlateID<-sprintf("10x10PL%03d", l)
DF10x10<-do.call(rbind,listOfTables)
write.csv(DF10x10, file="10x10dataframe.csv")
```


#Image-Plots of normalized viabilties
```{r}
colfunc <- colorRampPalette(c("darkgreen", "red"))
colviab<-colfunc(13)
pdf(file.path(outputdir,"10x10.pdf"), height=7, width=10)
unique(DF10x10$BaseDrug )
unique(DF10x10$PatientID )
for(pat in as.character(unique(DF10x10$PatientID ))){
  df4focus<-filter(DF10x10, PatientID==pat) 
    for(bdrug in as.character(unique(df4focus$BaseDrug ))){
      dftmp<-filter(df4focus, BaseDrug==bdrug)
      dftmp<-dftmp[dftmp$CombiDrug!="DM",]
      dftmp$base<-paste(dftmp$BaseDrug, dftmp$concBaseDrug, sep="_")
      dftmp$combi<-paste(dftmp$CombiDrug, dftmp$concCombiDrug, sep="_")
      dftmp<-select(dftmp, combi, base, normalizedValueDMSOControl)
      levelBase<-unique(dftmp$base)
      levelCombi<-unique(dftmp$combi)
     gg<- ggplot(dftmp, aes(x=factor(base,levels=levelBase), y=factor(combi, levels=levelCombi), fill=normalizedValueDMSOControl))+
        geom_tile()+ggtitle(paste(bdrug, pat))+scale_fill_gradient(limits=c(0,1.4), breaks=seq(0,1.2,0.2))
     print(gg)
    }  }
dev.off()
```



#Calculate IC_x
To determine for each drug and combinationt the IX_x concentrations fit dose-response curves across all patients.

??????

2d fit to get IC50 for each combination
how to calculate isobologram?
```{r, eval=F}
library(drc)
#base drug only
for(pat in unique(DF10x10$PatientID)){
for( drB in (setdiff(unique(DF10x10$BaseDrug), "DM"))){
if(nrow(filter(DF10x10, BaseDrug==drB & PatientID==pat ))>0){
dftmp<-DF10x10%>% filter(CombiDrug=='Ibrutinib'&concCvalue==0 & BaseDrug==drB & PatientID==pat)
m<-drm(normalizedValueDMSOControl ~ concBvalue, data=dftmp,fct = LL.4())
# print(summary(m))
# plot(m, main=paste(drB, pat))
newx=data.frame(concBvalue=seq(0,max(dftmp$concBvalue), length.out = 10000))
pred<-predict(m, newx)
IC50<-newx[which.min(abs(pred-0.5)),]
IC90<-newx[which.min(abs(pred-0.9)),]
IC70<-newx[which.min(abs(pred-0.7)),]
ICValuebase<-c(IC50=IC50, IC70=IC70, IC90=IC90)


#combi drug only in the corresponding 
dftmp<-DF10x10%>% filter(CombiDrug=='Ibrutinib'&concBvalue==0& BaseDrug==drB & PatientID==pat)
m<-drm(normalizedValueDMSOControl ~ concCvalue, data=dftmp,fct = LL.4())
# print(summary(m))
# plot(m, main=paste("Ibrutinib on plate", drB, pat))
newx=data.frame(concCvalue=seq(0,max(dftmp$concCvalue), length.out = 10000))
pred<-predict(m, newx)
IC50<-newx[which.min(abs(pred-0.5)),]
IC90<-newx[which.min(abs(pred-0.9)),]
IC70<-newx[which.min(abs(pred-0.7)),]
ICValueIbru<-c(IC50=IC50, IC70=IC70, IC90=IC90)

#IC values of combinations
dfcombi<-DF10x10%>% filter(CombiDrug=='Ibrutinib'& BaseDrug==drB & concBvalue!=0 & concCvalue!=0 & PatientID==pat)
plot(c(0,ICValueIbru["IC50"]), c(ICValuebase["IC50"],0), main=paste(drB, pat), type='l', xlab="concIbru", ylab="concBase")

#which way around (fit concB for conC fixed or otherwise, fit 2d?)
for(concB in unique(dfcombi$concBvalue)){
dftmp<-DF10x10%>% filter(CombiDrug=='Ibrutinib'&concBvalue==concB& BaseDrug==drB  & PatientID==pat)
if(any(dftmp$normalizedValueDMSOControl>0.5)){
m<-drm(normalizedValueDMSOControl ~ concCvalue, data=dftmp,fct = LL.4())
# print(summary(m))
# plot(m, main=paste("Ibrutinib on plate", drB, pat))
newx=data.frame(concCvalue=seq(0,max(dftmp$concCvalue), length.out = 10000))
pred<-predict(m, newx)
IC50<-newx[which.min(abs(pred-0.5)),]
IC90<-newx[which.min(abs(pred-0.9)),]
IC70<-newx[which.min(abs(pred-0.7)),]
ICValuejoint<-data.frame(concC=c(IC50=IC50, IC70=IC70, IC90=IC90), concB=concB)
print(ICValuejoint)
points(ICValuejoint["IC50",])
}
}
#which way around (fit concB for conC fixed or otherwise, fit 2d?)
for(concC in unique(dfcombi$concCvalue)){
dftmp<-DF10x10%>% filter(CombiDrug=='Ibrutinib'&concCvalue==concC& BaseDrug==drB & PatientID==pat)
if(any(dftmp$normalizedValueDMSOControl>0.5)){
m<-drm(normalizedValueDMSOControl ~ concBvalue, data=dftmp,fct = LL.4())
# print(summary(m))
# plot(m, main=paste("Ibrutinib on plate", drB, pat))
newx=data.frame(concBvalue=seq(0,max(dftmp$concCvalue), length.out = 10000))
pred<-predict(m, newx)
IC50<-newx[which.min(abs(pred-0.5)),]
IC90<-newx[which.min(abs(pred-0.9)),]
IC70<-newx[which.min(abs(pred-0.7)),]
ICValuejoint<-data.frame(concC=concC, concB=c(IC50=IC50, IC70=IC70, IC90=IC90))
print(ICValuejoint)
points(ICValuejoint["IC50",], pch=3)
}
}
}
}
}
```

#Isobologram
Concentrations required for given effect (e.g., IC50) are determined for drug B (ICx, B) and drug C (ICx, C). A line connecting (ICx, A, 0) and (0, ICx, B) is the line of additivity. Then, the concentrations of A and B contained in combination that provide the same effect, denoted as (CA, x, CB, x), are placed in the same plot. Synergy, additivity, or antagonism is indicated when (CA, x, CB, x) is located below, on, or above the line, respectively.

```{r}
```



#CI Loewe
$CI=\frac{C_{A,x}}{IC_{x,A}}+\frac{C_{B,x}}{IC_{x,B}}$
