---
title: "CLL drug combinations ex-vivo: Analyse effect of base drugs alone"
author: "Britta Velten"
date: "31 August 2018"
output:
  BiocStyle::html_document:
    toc: true
    toc_depth: 3
params: 
    today: 180905
---

#Introduction

This file analyses the data from the combination screen for combination with DMSO. (Effect of base drug alone)

```{r, echo=F}
library(knitr )
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

```{r, echo=F}
library(ggplot2)
library(RColorBrewer)
library(Biobase)
library(abind)
library(grid)
library(gtable)
library(reshape2)
library(gridExtra)
require(pracma)
library(reshape2)
library(dplyr)
library(pheatmap)
library(magrittr)
library(lattice)
library(tidyverse)
options(stringsAsFactors = FALSE)
```

```{r}
# setwd("~/Documents/cll/MarinaDrugComb/DrugCombi_Code4Ms/")
# datadir <- "~/Documents/cll/MarinaDrugComb/rawData"
datadir <- "/Volumes/huber/projects/nct/cll/RawData/DrugScreens/Marina_CombiScreen/"
outdir <- "out"
today <- params$today
figdir <- paste0("figuresBaseAlone", today, "/")
if(!dir.exists(figdir)) dir.create(figdir)
knitr::opts_chunk$set(dev = c("png", "pdf"), fig.path = figdir)
```


# Data Import:
Load processed Data from DrugCombi_QC.Rmd and DrugCombi_AddOmics (moelcular data on patient samples)
```{r}
load(file.path(outdir, paste0("MarinaCombiDataAfterQC_",today,".RData")))
load(file.path(outdir, paste0("OmicsData",today,".RData")))
```

```{r}
# discard outliers for further analyses, only use averaged values for replicates
df4ana <- df4anaAvreplicatesRmOutliers
rm(df4anaAvreplicatesRmOutliers, df4anaAvreplicates)
```

# Construct dataframe for base effect alone
Note that each experiment consists of 3 combination partner plus DMSO. Therefore, all patient samples with more than 3 combinaiton partners will have several effect values of the DMSO plate from each experiment. In the following we will average those values. For the combination studies comparisons are made within the experiments taken the value from the corresponding DMSO plate.

```{r}
df_effectB <- df4ana %>% select(starts_with("BDrug"), PatientID, effectB)

# this contains multiple values for most patient, drug, concentration pairs:
nrow(df_effectB)
52*32*5
df <- df_effectB %>%
  group_by(BDrugName, BDrugID, BDrugConcId,BDrugConc,PatientID) %>%
  summarize(n_effectB = length(effectB)) 
table(df$n_effectB)

#summarize by mean
df_effectB %<>%  group_by(BDrugName, BDrugID, BDrugConcId,BDrugConc,PatientID) %>%
  summarize(effectB = mean(effectB))%>%
  ungroup()
nrow(df_effectB)
```

# Test for assciations of base effect to mutations
```{r}
df_effectB_vs_muts <- left_join(df_effectB,
                                gather(dfMuts4testing, key="mutation", value="status", -PatientID), by="PatientID")
df_effectB_vs_muts <- df_effectB_vs_muts[!duplicated(df_effectB_vs_muts),] #remove duplicates, as info is contained in each combi durg
df_pvals_effectB_vs_muts <- group_by(df_effectB_vs_muts, BDrugName, BDrugID, BDrugConc, BDrugConcId, mutation) %>%
  summarize(pval = t.test(effectB ~ status, var.equal= TRUE)$p.value,
            mean.diff = diff(t.test(effectB ~ status, var.equal= TRUE)$estimate))
df_pvals_effectB_vs_muts$pval.adj <- p.adjust(df_pvals_effectB_vs_muts$pval, method="BH")
hist(df_pvals_effectB_vs_muts$pval)
```


```{r}
# FDR threshold
alpha <- 0.05
```

## Volcanoe plot
```{r volcanoe_base, fig.height=7, fig.width=7}
# take only most significant concentration for plotting (FDR control is on level of single concentrations)
df_pvals_max <- df_pvals_effectB_vs_muts %>% group_by(BDrugName, BDrugID, mutation) %>%
  summarize(idx = which.min(pval), pval = pval[idx], pval.adj= pval.adj[idx], mean.diff = mean.diff[idx], conc = BDrugConcId[idx])
df_pvals_max %<>% mutate(sig = pval.adj<alpha, col = mean.diff>0)

ggplot(df_pvals_max, aes(x=mean.diff, y=-log10(pval.adj), col=col)) +
  geom_point(alpha=0.7) + ggrepel::geom_label_repel(aes(label=ifelse(sig, BDrugName, ""))) +
  scale_color_manual(values= c("deeppink", "navy")) + facet_wrap(~mutation)+ guides(col=FALSE)
```

## List of significant pharmacogenomic interactions
```{r}
df_sign <- df_pvals_effectB_vs_muts %>% filter(pval.adj < alpha)
df_nconc <- df_sign %>% group_by(BDrugName, mutation) %>% summarize(nconc =n())

# TP53
df_signTP53 <- df_sign %>% filter(mutation == "TP53")
paste(unique(df_signTP53$BDrugName), collapse = ", ")

# IGHV
df_signIGHV <- df_sign %>% filter(mutation == "IGHV")
paste(unique(df_signIGHV$BDrugName), collapse = ", ")
```

## Boxplots
Draw boxplots for TP53 and IGHV including all drugs that have significant concentrations at an FDR of 1%.
```{r boxplot_base_TP53, fig.width=9, fig.height=4}
df_effectB_vs_muts %>% filter(mutation == "TP53", BDrugName %in% df_signTP53$BDrugName) %>% 
  mutate(status = ifelse(status ==0, "wt","mut")) %>%
  ggplot(aes(x=BDrugConcId, y=effectB)) + geom_boxplot(aes(fill=factor(status))) + facet_wrap(~BDrugName) +
  ggpubr::stat_compare_means(method = "t.test", aes(group=status, label =  ..p.signif..), hide.ns = TRUE) +
  guides(fill = guide_legend(title="TP53")) + xlab("Concentration") +ylab("viability") +theme_bw()

```

```{r, boxplot_base_IGHV, fig.width=9, fig.height=9}
df_effectB_vs_muts %>% filter(mutation == "IGHV", BDrugName %in% df_signIGHV$BDrugName) %>% 
    mutate(status = ifelse(status ==0, "U-CLL","M-CLL")) %>%
  ggplot(aes(x=BDrugConcId, y=effectB)) + geom_boxplot(aes(fill=factor(status))) + facet_wrap(~BDrugName) +
  ggpubr::stat_compare_means(method = "t.test", aes(group=status, label =  ..p.signif..), hide.ns = TRUE) +
  guides(fill = guide_legend(title="IGHV"))+ xlab("Concentration") +ylab("viability")+theme_bw()
```


## Heatmap on base viability values
```{r}
df_effectB %<>% mutate(BDrugNameConc = paste(BDrugName, BDrugConcId, sep="_"))
effectB_mat <- df_effectB %>% 
  select(BDrugNameConc, PatientID, effectB)  %>%
  spread(key=BDrugNameConc, value=effectB) %>% 
  column_to_rownames("PatientID") %>%
  as.matrix()
```

## Base-DrugxDrug
```{r heatmap_base_drug_drug}
pheatmap(cor(effectB_mat, use = "complete.obs"),
         show_rownames = FALSE,treeheight_row = 10, treeheight_col = 10)
```


Average over concentrations
```{r heatmap_base_drug_drug_av, fig.height=7}
effectB_mat_av <- df_effectB %>% 
  select(BDrugName, PatientID, effectB)  %>%
  group_by(BDrugName,PatientID) %>%
  summarise(effectB = mean(effectB, na.rm=TRUE)) %>%
  ungroup() %>%
  spread(key=BDrugName, value=effectB) %>% 
  column_to_rownames("PatientID") %>%
  as.matrix()
pheatmap(cor(effectB_mat_av, use = "complete.obs"), show_rownames = FALSE,treeheight_row = 10, treeheight_col = 10)
```

##PatientxDrugs
```{r heatmap_base_patient_drug}
# outlying values cut off at 1.4
effectB_mat[is.na(effectB_mat)] <- 1.4

# add genetic background
dfanno <- dfMuts4testing %>% column_to_rownames("PatientID") %>% select(which(colSums(.) > 10))
dfanno <- as.data.frame(ifelse(dfanno==1, "mut", "wt"))

# set colors for mutation status
cols_mut        <- c("white", "black")
names(cols_mut) <- c("wt", "mut")
anno_colors <- lapply(colnames(dfanno), function(x) cols_mut)
names(anno_colors) <- colnames(dfanno)

 pheatmap(effectB_mat, na_col="gray", clustering_distance_rows="euclidean",
         clustering_distance_cols = "euclidean",
         color = colorRampPalette(rev(brewer.pal(n = 7, name ="RdYlBu")))(140),
         breaks=seq(0,filter_th,0.01), show_rownames = FALSE, show_colnames = FALSE,
         treeheight_row = 10, treeheight_col = 10, annotation_row = dfanno, annotation_colors =anno_colors, annotation_legend = FALSE)
```

Average over concentrations
```{r heatmap_base_patient_drug_av}
effectB_mat_av <- df_effectB %>% 
  select(BDrugName, PatientID, effectB)  %>%
  group_by(BDrugName,PatientID) %>%
  summarise(effectB = mean(effectB, na.rm=TRUE)) %>%
  ungroup() %>%
  spread(key=BDrugName, value=effectB) %>% 
  column_to_rownames("PatientID") %>%
  as.matrix()

# add genetic background
dfanno <- dfMuts4testing %>% column_to_rownames("PatientID") %>% select(which(colSums(.) > 10))
dfanno <- as.data.frame(ifelse(dfanno==1, "mut", "wt"))

# set colors for mutation status
cols_mut        <- c("white", "black")
names(cols_mut) <- c("wt", "mut")
anno_colors <- lapply(colnames(dfanno), function(x) cols_mut)
names(anno_colors) <- colnames(dfanno)

 pheatmap(effectB_mat_av, na_col="gray", clustering_distance_rows="v",
         clustering_distance_cols = "euclidean",
         color = colorRampPalette(rev(brewer.pal(n = 7, name ="RdYlBu")))(140),
         breaks=seq(0,filter_th,0.01), show_rownames = FALSE, show_colnames = TRUE,
         treeheight_row = 10, treeheight_col = 10, annotation_row = dfanno, annotation_colors =anno_colors, annotation_legend = FALSE)
```


# Single correlation values
```{r}
# navitoclax - venetoclax
cor(as.numeric(effectB_mat[,grep("Navitoclax", colnames(effectB_mat))]), as.numeric(effectB_mat[,grep("Venetoclax", colnames(effectB_mat))]))

cor(as.numeric(effectB_mat[,grep("YM155", colnames(effectB_mat))]), as.numeric(effectB_mat[,grep("SNS-032", colnames(effectB_mat))]))

```

#SessionInfo
```{r}
sessionInfo()
```

