---
title: "CLL drug combinations ex-vivo: 10x10 screens"
author: "Britta Velten"
date: "4 September 2018"
output:
  BiocStyle::html_document:
    toc: true
    toc_depth: 3
params: 
    today: 180903
---

#Introduction

Takes raw data files for 10x10 screen.

```{r, echo=F}
library(knitr )
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
options(stringAsFactors = FALSE)
```

```{r, echo=F}
library(ggplot2)
library(RColorBrewer)
library(Biobase)
library(abind)
library(grid)
library(gtable)
library(reshape2)
library(gridExtra)
require(pracma)
library(reshape2)
library(dplyr)
library(pheatmap)
library(magrittr)
library(lattice)
library(tidyverse)
library(xlsx)
source("plot_utils.R")
```

```{r}
# setwd("~/Documents/cll/MarinaDrugComb/DrugCombi_Code4Ms/")
# datadir <- "~/Documents/cll/MarinaDrugComb/rawData"
datadir <- "/Volumes/huber/projects/nct/cll/RawData/DrugScreens/Marina_CombiScreen/"
outdir = "out"
today <- params$today
knitr::opts_chunk$set(dev = c("png", "pdf"), fig.path = paste0("figures10x10", today, "/"))
```


```{r}
datadir10x10 <- file.path(datadir,"10x10")
```


# Get plate eet-up
From Prtocol get set-up of the plate (i.e. which substance in which concentration in which well)
```{r}
FileSetUp10x10 <- as.matrix(read.xlsx(file.path(datadir10x10,"Protokoll 10 x 10 Validierung Drugscreen2.xlsx"),2))

# get set-up for base/test drugs
  tableIdxBase <- which(FileSetUp10x10=="A", arr.ind = T)[1,]
  SetUpBase <- FileSetUp10x10[tableIdxBase[1]+0:15,tableIdxBase[2]+1:24]
  rownames(SetUpBase) <- FileSetUp10x10[tableIdxBase[1]+0:15,tableIdxBase[2]]
  colnames(SetUpBase) <- FileSetUp10x10[tableIdxBase[1]-1,tableIdxBase[2]+1:24]
  SetUpBase <- rownames_to_column(as.data.frame(SetUpBase), var = "WellLetter") %>%
    gather(key="WellNo", value = "base", -WellLetter)
  
  # get concentration values per column of plate
  conc_base_df <- tibble(WellNo = FileSetUp10x10[tableIdxBase[1]-1,tableIdxBase[2]+1:24],
                      base_conc = FileSetUp10x10[tableIdxBase[1]-2,tableIdxBase[2]+1:24])
  
  SetUpBase %<>% left_join(conc_base_df, by = "WellNo")
  SetUpBase %<>% mutate(base_conc = ifelse(base=="DM", NA, base_conc))
  SetUpBase <- filter(SetUpBase, !is.na(base))
  
  
# get set-up for ibrutinib
  tableIdxIbr <- which(FileSetUp10x10=="A", arr.ind = T)[2,]
  SetUpIbr <- FileSetUp10x10[tableIdxIbr[1]+0:15,tableIdxIbr[2]+1:24]
  rownames(SetUpIbr) <- FileSetUp10x10[tableIdxIbr[1]+0:15,tableIdxIbr[2]]
  colnames(SetUpIbr) <- FileSetUp10x10[tableIdxIbr[1]-1,tableIdxIbr[2]+1:24]
  SetUpIbr <- rownames_to_column(as.data.frame(SetUpIbr), var = "WellLetter") %>%
    gather(key="WellNo", value = "combi", -WellLetter)
  
  # get concentration values per row of plate
  conc_combi_df <- tibble(WellLetter = FileSetUp10x10[tableIdxIbr[1]+0:15,tableIdxIbr[2]],
                          combi_conc = FileSetUp10x10[tableIdxIbr[1]+(0:15),tableIdxIbr[2]+25])
  conc_combi_df %<>% mutate(combi_conc = ifelse(combi_conc == "Concentration (Synergie-Substanz)", NA, combi_conc))

  SetUpIbr %<>% left_join(conc_combi_df, by = "WellLetter")
  
  # set conc for DM to NA
  SetUpIbr %<>% mutate(combi_conc = ifelse(!is.na(combi), ifelse(combi=="DM", NA, combi_conc), combi_conc))

  #fill in NAs with a conc value by Ibrutinib
  SetUpIbr %<>% mutate(combi = ifelse(combi_conc %in% paste0("c", 1:10), "Ibrutinib", combi))
  SetUpIbr <- filter(SetUpIbr, !is.na(combi))
                       
  # fill in row with changing values
  SetUpIbr %<>% mutate(combi_conc = ifelse(grepl("Ibru c",combi), sub("Ibru ", "", combi), combi_conc))
  SetUpIbr %<>% mutate(combi = ifelse(grepl("Ibru c",combi), "Ibrutinib", combi))

  # merge 
  SetUp <- full_join(SetUpIbr, SetUpBase, by=c("WellLetter", "WellNo"))
  SetUp %<>% mutate(WellLetter = tolower(WellLetter))
```


# Read in values from 10x10 plates
```{r}
import10x10Data <- function(filename, SetUpBase, SetUpIbr){
  print(filename)
  data <- read.xlsx2(file.path(datadir10x10,filename),1, header = F, startRow=3)
  
  #info from filename
  NameComp <- strsplit(filename, "_")[[1]]
  patID <- NameComp[grep("P0",NameComp)] 
  TestDrugA <- NameComp[3] 
  TestDrugB <- sub(".xlsx","", NameComp[4])
  
  #table from excel file (ignore meta data about screen)
  tableIdx <- which(data=="<>", arr.ind = T)
  stopifnot(nrow(tableIdx)==1)
  plateValues <- data[tableIdx[1]+1:11,tableIdx[2]+1:24]
  colnames(plateValues) <- 1:ncol(plateValues)
  rownames(plateValues) <- letters[2 + 1:nrow(plateValues)]  #labels start with C
  
  # turn into data.frame
  df <- rownames_to_column(plateValues, var = "WellLetter")
  df %<>% gather(key="WellNo", value="rawValue", -WellLetter)
  df %<>% mutate(rawValue = as.numeric(rawValue))
  
  # add setup info
  df %<>% left_join(SetUp,  by=c("WellLetter", "WellNo"))
  
  # fill in test drug names: test drug 1 is first of file name
  df %<>% mutate(base = ifelse(base =="DM", "DM", ifelse(base==1,TestDrugA, TestDrugB)))
  df %<>% rename(BaseDrug = base)
  df %<>% rename(CombiDrug = combi)

  # normalized by pure DMSO median
  dfDMSO <- filter(df, BaseDrug=="DM" , CombiDrug=="DM")
  df %<>% mutate(pureDMSOMedian = median(dfDMSO$rawValue))
  df %<>% mutate(pureDMSOMean = mean(dfDMSO$rawValue))
  df %<>% mutate(normalizedValue = rawValue/pureDMSOMedian)

  # get table of concentrations (for Ibrutinib, drug A and drug B)
  # (bug in reaad.xlsx but dilution factor is always 2 and c10 is zero, replace by hand)
  conctableIdx <- which(data=="µM"| data=="nM", arr.ind = T)
  concArray <- sapply(1:3, function(i) {
    concvec <- rep(0,10)
    names(concvec) <- paste("c", 1:10, sep="")
    concvec[1] <- as.numeric(as.character(data[conctableIdx[i,1]+0,conctableIdx[i,2]+1]))
    if(data[conctableIdx[i,1],conctableIdx[i,2]]=="nM")   concvec[1] <- concvec[1]/1000
    for(j in 2:9) concvec[j] <- concvec[j-1]/2
    concvec
  })
  colnames(concArray) <- data[conctableIdx-c(1,1)]
  concArray <- cbind(concArray,DM=rep(NA,10))
  colnames(concArray)[colnames(concArray)=="CAL-101"] <- "CAL101"
  colnames(concArray)[colnames(concArray)=="Abt-199"] <- "Abt199"
  colnames(concArray)[colnames(concArray)=="ABT-263"] <- "ABT263"
  concValueDf <- concArray %>% as.data.frame() %>% 
    rownames_to_column(var = "concId") %>% gather(key="drug", value = "concValue", -concId)
  
  # add true conc values to df
  df %<>% left_join(concValueDf, by = c("CombiDrug" = "drug", "combi_conc" = "concId"))
  df %<>% rename(concCvalue = concValue)

  df %<>% left_join(concValueDf, by = c("BaseDrug" = "drug", "base_conc" = "concId"))
  df %<>% rename(concBvalue = concValue)
  df %<>% mutate(PatientID = patID)

  return(df)
  }
```

Different from Marina normalized values by median instead of mean (as done in other combi screens).
```{r}
allFiles <- setdiff(list.files(file.path(datadir10x10), "*.xlsx"), list.files(file.path(datadir10x10),"Protokoll"))
listOfTables <- lapply(allFiles, function(file) import10x10Data(file, SetUpBase, SetUpIbr))
for(l in 1:length(listOfTables)) listOfTables[[l]]$PlateID <- sprintf("10x10PL%03d", l)
df10x10 <- do.call(rbind,listOfTables)
```

```{r, echo=FALSE}
# df10x10_old <- read.csv("~/Documents/CLL/MarinaDrugComb/Analysis/ProcessedDataTables/10x10dataframe.csv")
# df10x10_old %<>% rename(normalizedValue = normalizedValueDMSOControl,
#                     base_conc = concBaseDrug,
#                     combi_conc = concCombiDrug)
# df10x10$WellNo <- as.numeric(df10x10$WellNo)
# samecols <- intersect(colnames(df10x10), colnames(df10x10_old))
# 
# df <- full_join(df10x10_old, df10x10, by=c("CombiDrug", "BaseDrug", "base_conc", "combi_conc", "PatientID", "WellNo", "WellLetter"))
# all(df$rawValue.x == df$rawValue.y)
# max(abs(df$normalizedValue.x- df$normalizedValue.y))
# max(abs(df$pureDMSOMean.x- df$pureDMSOMean.y))
# max(abs(df$pureDMSOMedian.x- df$pureDMSOMedian.y))
```

```{r}
# use nice drug names
df10x10 %<>% mutate(BaseDrugName = ifelse(BaseDrug == "CAL101", "Idelalisib", 
                           ifelse(BaseDrug == "Abt199", "Venetoclax",
                                  ifelse(BaseDrug == "ABT263","Navitoclax", BaseDrug))))
# data overview
df10x10 %>% group_by(BaseDrugName, CombiDrug) %>%
  summarize(nPats = length(unique(PatientID)), pats = paste(unique(PatientID), collapse =", "))
```

```{r}
save(df10x10, file = file.path(outdir, paste0("df10x10_",today,".Rdata")))
```


# Tile-Plots of normalized viabilties

```{r}
knitr::opts_chunk$set(fig.width=8)
```

## Navitoclax
```{r 10x10_Navitoclax_P0016}
plotTiles10x10(df10x10, "Navitoclax", "P0016")
```
```{r 10x10_Navitoclax_P0347}
plotTiles10x10(df10x10, "Navitoclax", "P0347")
```
```{r 10x10_Navitoclax_P0445}
plotTiles10x10(df10x10, "Navitoclax", "P0445")
```
```{r 10x10_Navitoclax_P0309}
plotTiles10x10(df10x10, "Navitoclax", "P0309")
```
```{r 10x10_Navitoclax_P0369}
plotTiles10x10(df10x10, "Navitoclax", "P0369")
```

## Afatinib
```{r 10x10_Afatinib_P0010}
plotTiles10x10(df10x10, "Afatinib", "P0010")
```
```{r 10x10_Afatinib_P0050}
plotTiles10x10(df10x10, "Afatinib", "P0050")
```
```{r 10x10_Afatinib_P0445}
plotTiles10x10(df10x10, "Afatinib", "P0069")
```
```{r 10x10_Afatinib_P0480}
plotTiles10x10(df10x10, "Afatinib", "P0480")
```
```{r 10x10_Afatinib_P0484}
plotTiles10x10(df10x10, "Afatinib", "P0484")
```

## Venetoclax
```{r 10x10_Venetoclax_P0016}
plotTiles10x10(df10x10, "Venetoclax", "P0016")
```
```{r 10x10_Venetoclax_P0347}
plotTiles10x10(df10x10, "Venetoclax", "P0347")
```
```{r 10x10_Venetoclax_P0445}
plotTiles10x10(df10x10, "Venetoclax", "P0445")
```


```{r}
sessionInfo()
```


```{r}
knitr::opts_chunk$set(eval = FALSE, echo = FALSE)
```

#Calculate IC_x
To determine for each drug and combinationt the IX_x concentrations fit dose-response curves across all patients.

??????

2d fit to get IC50 for each combination
how to calculate isobologram?
```{r, eval=F}
library(drc)
#base drug only
for(pat in unique(DF10x10$PatientID)){
for( drB in (setdiff(unique(DF10x10$BaseDrug), "DM"))){
if(nrow(filter(DF10x10, BaseDrug==drB & PatientID==pat ))>0){
dftmp<-DF10x10%>% filter(CombiDrug=='Ibrutinib'&concCvalue==0 & BaseDrug==drB & PatientID==pat)
m<-drm(normalizedValueDMSOControl ~ concBvalue, data=dftmp,fct = LL.4())
# print(summary(m))
# plot(m, main=paste(drB, pat))
newx=data.frame(concBvalue=seq(0,max(dftmp$concBvalue), length.out = 10000))
pred<-predict(m, newx)
IC50<-newx[which.min(abs(pred-0.5)),]
IC90<-newx[which.min(abs(pred-0.9)),]
IC70<-newx[which.min(abs(pred-0.7)),]
ICValuebase<-c(IC50=IC50, IC70=IC70, IC90=IC90)


#combi drug only in the corresponding 
dftmp<-DF10x10%>% filter(CombiDrug=='Ibrutinib'&concBvalue==0& BaseDrug==drB & PatientID==pat)
m<-drm(normalizedValueDMSOControl ~ concCvalue, data=dftmp,fct = LL.4())
# print(summary(m))
# plot(m, main=paste("Ibrutinib on plate", drB, pat))
newx=data.frame(concCvalue=seq(0,max(dftmp$concCvalue), length.out = 10000))
pred<-predict(m, newx)
IC50<-newx[which.min(abs(pred-0.5)),]
IC90<-newx[which.min(abs(pred-0.9)),]
IC70<-newx[which.min(abs(pred-0.7)),]
ICValueIbru<-c(IC50=IC50, IC70=IC70, IC90=IC90)

#IC values of combinations
dfcombi<-DF10x10%>% filter(CombiDrug=='Ibrutinib'& BaseDrug==drB & concBvalue!=0 & concCvalue!=0 & PatientID==pat)
plot(c(0,ICValueIbru["IC50"]), c(ICValuebase["IC50"],0), main=paste(drB, pat), type='l', xlab="concIbru", ylab="concBase")

#which way around (fit concB for conC fixed or otherwise, fit 2d?)
for(concB in unique(dfcombi$concBvalue)){
dftmp<-DF10x10%>% filter(CombiDrug=='Ibrutinib'&concBvalue==concB& BaseDrug==drB  & PatientID==pat)
if(any(dftmp$normalizedValueDMSOControl>0.5)){
m<-drm(normalizedValueDMSOControl ~ concCvalue, data=dftmp,fct = LL.4())
# print(summary(m))
# plot(m, main=paste("Ibrutinib on plate", drB, pat))
newx=data.frame(concCvalue=seq(0,max(dftmp$concCvalue), length.out = 10000))
pred<-predict(m, newx)
IC50<-newx[which.min(abs(pred-0.5)),]
IC90<-newx[which.min(abs(pred-0.9)),]
IC70<-newx[which.min(abs(pred-0.7)),]
ICValuejoint<-data.frame(concC=c(IC50=IC50, IC70=IC70, IC90=IC90), concB=concB)
print(ICValuejoint)
points(ICValuejoint["IC50",])
}
}
#which way around (fit concB for conC fixed or otherwise, fit 2d?)
for(concC in unique(dfcombi$concCvalue)){
dftmp<-DF10x10%>% filter(CombiDrug=='Ibrutinib'&concCvalue==concC& BaseDrug==drB & PatientID==pat)
if(any(dftmp$normalizedValueDMSOControl>0.5)){
m<-drm(normalizedValueDMSOControl ~ concBvalue, data=dftmp,fct = LL.4())
# print(summary(m))
# plot(m, main=paste("Ibrutinib on plate", drB, pat))
newx=data.frame(concBvalue=seq(0,max(dftmp$concCvalue), length.out = 10000))
pred<-predict(m, newx)
IC50<-newx[which.min(abs(pred-0.5)),]
IC90<-newx[which.min(abs(pred-0.9)),]
IC70<-newx[which.min(abs(pred-0.7)),]
ICValuejoint<-data.frame(concC=concC, concB=c(IC50=IC50, IC70=IC70, IC90=IC90))
print(ICValuejoint)
points(ICValuejoint["IC50",], pch=3)
}
}
}
}
}
```

#Isobologram
Concentrations required for given effect (e.g., IC50) are determined for drug B (ICx, B) and drug C (ICx, C). A line connecting (ICx, A, 0) and (0, ICx, B) is the line of additivity. Then, the concentrations of A and B contained in combination that provide the same effect, denoted as (CA, x, CB, x), are placed in the same plot. Synergy, additivity, or antagonism is indicated when (CA, x, CB, x) is located below, on, or above the line, respectively.

```{r}
```



#CI Loewe
$CI=\frac{C_{A,x}}{IC_{x,A}}+\frac{C_{B,x}}{IC_{x,B}}$
